### 2024-05-25 paper/figs/definitions.tex
some work on model diagram for paper

\usepackage{xcolor}

% Color definitions
\definecolor{customorange}{rgb}{0.94, 0.55, 0}
\definecolor{customteal}{rgb}{0.47, 0.52, 0.6}
\definecolor{customgreen}{rgb}{0.18, 0.62, 0.27}
\definecolor{fontcolor}{rgb}{0.204, 0.227, 0.251}

% Tikz shape definitions
\tikzstyle{nn-box} = [rectangle, rounded corners=0.5mm, minimum width=35mm, minimum height=10mm, text centered, draw=customteal, fill=customteal!30, text=fontcolor]
\tikzstyle{transformer-box} = [rectangle, rounded corners=0.5mm, minimum width=35mm, minimum height=10mm, text centered, draw=customgreen, fill=customgreen!30, text=fontcolor]
\tikzstyle{xgb-box} = [rectangle, rounded corners=0.5mm, minimum width=35mm, minimum height=10mm, text centered, draw=customorange, fill=customorange!30, text=fontcolor]
\tikzstyle{pca-box} = [rectangle, rounded corners=0.5mm, minimum width=20mm, minimum height=7mm, text centered, draw=customorange!70, fill=customorange!10, text=fontcolor]

% \tikzstyle{startstop} = [rectangle, rounded corners, minimum width=3cm, minimum height=1cm,text centered, draw=black, fill=red!30]
% \tikzstyle{process} = [rectangle, minimum width=3cm, minimum height=1cm, text centered, draw=black, fill=orange!30]
% \tikzstyle{arrow} = [thick,->,>=stealth]

### 2024-05-25 paper/figs/model-diagram.tex
some work on model diagram for paper

% Tikz shape definitions
\tikzstyle{nnbox} = [rectangle, rounded corners=0.5mm, minimum width=35mm, minimum height=10mm, text centered, draw=customteal, fill=customteal!30, text=fontcolor]
\tikzstyle{transformerbox} = [rectangle, rounded corners=0.5mm, minimum width=35mm, minimum height=10mm, text centered, draw=customgreen, fill=customgreen!30, text=fontcolor]
\tikzstyle{xgbbox} = [rectangle, rounded corners=0.5mm, minimum width=35mm, minimum height=10mm, text centered, draw=customorange, fill=customorange!30, text=fontcolor]
\tikzstyle{pcabox} = [rectangle, rounded corners=2mm, minimum width=20mm, minimum height=7mm, text centered, draw=customorange!70, fill=customorange!10, text=fontcolor!70]
\tikzstyle{databox} = [rectangle, minimum width=40mm, minimum height=12mm, text centered, draw=red!70, fill=red!10, text=fontcolor!70]
\tikzstyle{convnode} = [ellipse, minimum width=5mm, minimum height=5mm, fill=customteal!40, draw=customteal!80]
\tikzstyle{sarrowxgb} = [->, thick, draw=customorange!70]
\tikzstyle{sarrownn} = [->, thick, draw=customteal!70]
\tikzstyle{sarrowtr} = [->, thick, draw=customgreen!70]

\begin{tikzpicture}

% Data
\node (bioclim) [databox] {Bioclimatic Rasters};
\node (landsat) [databox, below of=bioclim, yshift=-7mm] {LandSat};
\node (metadata) [databox, below of=landsat, yshift=-7mm] {Metadata};
\node (sentinel) [databox, below of=metadata, yshift=-7mm] {Sentinel};

% PCAs
\node (pca0) [pcabox, right of=bioclim, xshift=30mm, yshift=30mm] {PCA$_0$}; 
\node (pca1) [pcabox, below of=pca0] {PCA$_1$}; 
\node (vertellipsis0) [below of=pca1, yshift=3mm] {\Large{\textcolor{fontcolor!70}{$\cdot$}}};
\node (vertellipsis1) [below of=vertellipsis0, yshift=9mm] {\Large{\textcolor{fontcolor!70}{$\cdot$}}};
\node (vertellipsis2) [below of=vertellipsis1, yshift=9mm] {\Large{\textcolor{fontcolor!70}{$\cdot$}}};
\node (pca2) [pcabox, below of=vertellipsis2, yshift=3mm] {PCA$_n$}; 
\draw[sarrowxgb] (bioclim.east) .. controls +(right:5mm) and +(left:8mm) .. (pca0.west);
\draw[sarrowxgb] (bioclim.east) .. controls +(right:5mm) and +(left:8mm) .. (pca1.west);
\draw[sarrowxgb] (bioclim.east) .. controls +(right:5mm) and +(left:8mm) .. (pca2.west);

% XGB
\node (xgb) [xgbbox, right of=pca0, xshift=30mm, yshift=-15mm] {XGBoostRegressor};
\draw[sarrowxgb] (pca0.east) .. controls +(right:5mm) and +(left:8mm) .. (xgb.west);
\draw[sarrowxgb] (pca1.east) .. controls +(right:5mm) and +(left:8mm) .. (xgb.west);
\draw[sarrowxgb] (pca2.east) .. controls +(right:5mm) and +(left:8mm) .. (xgb.west);
\draw[sarrowxgb] (landsat.east) .. controls +(right:35mm) and +(left:10mm) .. (xgb.west);
\draw[sarrowxgb] (metadata.east) .. controls +(right:35mm) and +(left:10mm) .. (xgb.west);

% ResNet
\node (nn0) [nnbox, below of=xgb, yshift=-15mm] {ResNet18};
\node (nn1) [nnbox, below of=nn0, yshift=-3mm] {ResNet18};
\node (nn2) [nnbox, below of=nn1, yshift=-3mm] {ResNet18};
\node (conv) [convnode, right of=nn1, xshift=20mm] {};
\draw[sarrownn] (bioclim.east) .. controls +(right:25mm) and +(left:20mm) .. (nn0.west);
\draw[sarrownn] (landsat.east) .. controls +(right:25mm) and +(left:20mm) .. (nn1.west);
\draw[sarrownn] (sentinel.east) .. controls +(right:25mm) and +(left:20mm) .. (nn2.west);
\draw[sarrownn] (nn0.east) .. controls +(right:7mm) and +(left:7mm) .. (conv.west);
\draw[sarrownn] (nn1.east) .. controls +(right:7mm) and +(left:7mm) .. (conv.west);
\draw[sarrownn] (nn2.east) .. controls +(right:7mm) and +(left:7mm) .. (conv.west);

% Transformer
\node (transformer) [transformerbox, below of=nn2, yshift=-15mm] {Transformer};
\draw[sarrowtr] (bioclim.east) .. controls +(right:25mm) and +(left:20mm) .. (transformer.west);
\draw[sarrowtr] (landsat.east) .. controls +(right:25mm) and +(left:20mm) .. (transformer.west);
\draw[sarrowtr] (metadata.east) .. controls +(right:25mm) and +(left:20mm) .. (transformer.west);
\draw[sarrowtr] (sentinel.east) .. controls +(right:25mm) and +(left:20mm) .. (transformer.west);

\end{tikzpicture}

### 2024-05-25 paper/main.tex
some work on model diagram for paper

\documentclass{article}
% \usepackage[utf8]{inputenc}
\usepackage{xcolor}
\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows}

% Color definitions
\definecolor{customorange}{rgb}{0.94, 0.55, 0}
\definecolor{customteal}{rgb}{0.47, 0.52, 0.6}
\definecolor{customgreen}{rgb}{0.18, 0.62, 0.27}
\definecolor{fontcolor}{rgb}{0.204, 0.227, 0.251}

\author{
    Rawlings, Darren\\
    \texttt{email@domain.loc}      % If needed
    \and 
    Chopard, Timothy\\
    \texttt{email@domain.loc}
}

\title{An Inspiring and Groundbreaking Look at Plants}

\begin{document}
    \maketitle
    \include{sections/01-intro}
    \include{figs/model-diagram}
\end{document}

### 2024-05-25 paper/main.tex

@@
-\documentclass{article}
-% \usepackage[utf8]{inputenc}
-\usepackage{xcolor}
-\usepackage{tikz}
-\usetikzlibrary{shapes.geometric, arrows}
+%% The first command in your LaTeX source must be the \documentclass command.
+%%
+%% Options:
+%% twocolumn : Two column layout.
+%% hf: enable header and footer.
+\documentclass[
+% twocolumn,
+% hf,
+]{ceurart}
+
+%%
+%% One can fix some overfulls
+\sloppy
+
+%%
+%% Minted listings support 
+%% Need pygment <http://pygments.org/> <http://pypi.python.org/pypi/Pygments>
+\usepackage{minted}
+%% auto break lines
+\setminted{breaklines=true}
 
@@
 \definecolor{fontcolor}{rgb}{0.204, 0.227, 0.251}
+%%
+%% end of the preamble, start of the body of the document source.
+\begin{document}
 
-\author{
-    Rawlings, Darren\\
-    \texttt{email@domain.loc}      % If needed
-    \and 
-    Chopard, Timothy\\
-    \texttt{email@domain.loc}
-}
+%%
+%% Rights management information.
+%% CC-BY is default license.
+\copyrightyear{2022}
+\copyrightclause{Copyright for this paper by its authors.
+  Use permitted under Creative Commons License Attribution 4.0
+  International (CC BY 4.0).}
 
-\title{An Inspiring and Groundbreaking Look at Plants}
+%%
+%% This command is for the conference information
+\conference{Woodstock'22: Symposium on the irreproducible science,
+  June 07--11, 2022, Woodstock, NY}
 
-\begin{document}
+%%
+%% The "title" command
+\title{A better way to format your document for CEUR-WS}
+
+\tnotemark[1]
+\tnotetext[1]{You can use this document as the template for preparing your
+  publication. We recommend using the latest version of the ceurart style.}
+
+%%
+%% The "author" command and its associated commands are used to define
+%% the authors and their affiliations.
+\author[1,2]{Dmitry S. Kulyabov}[%
+orcid=0000-0002-0877-7063,
+email=kulyabov-ds@rudn.ru,
+url=https://yamadharma.github.io/,
+]
+\cormark[1]
+\fnmark[1]
+\address[1]{Peoples' Friendship University of Russia (RUDN University),
+  6 Miklukho-Maklaya St, Moscow, 117198, Russian Federation}
+\address[2]{Joint Institute for Nuclear Research,
+  6 Joliot-Curie, Dubna, Moscow region, 141980, Russian Federation}
+
+\author[3]{Ilaria Tiddi}[%
+orcid=0000-0001-7116-9338,
+email=i.tiddi@vu.nl,
+url=https://kmitd.github.io/ilaria/,
+]
+\fnmark[1]
+\address[3]{Vrije Universiteit Amsterdam, De Boelelaan 1105, 1081 HV Amsterdam, The Netherlands}
+
+\author[4]{Manfred Jeusfeld}[%
+orcid=0000-0002-9421-8566,
+email=Manfred.Jeusfeld@acm.org,
+url=http://conceptbase.sourceforge.net/mjf/,
+]
+\fnmark[1]
+\address[4]{University of Skövde, Högskolevägen 1, 541 28 Skövde, Sweden}
+
+%% Footnotes
+\cortext[1]{Corresponding author.}
+\fntext[1]{These authors contributed equally.}
+
+%%
+%% The abstract is a short summary of the work to be presented in the
+%% article.
+\begin{abstract}
+  A clear and well-documented \LaTeX{} document is presented as an
+  article formatted for publication by CEUR-WS in a conference
+  proceedings. Based on the ``ceurart'' document class, this article
+  presents and explains many of the common variations, as well as many
+  of the formatting elements an author may use in the preparation of
+  the documentation of their work.
+\end{abstract}
+
+%%
+%% Keywords. The author(s) should pick words that accurately describe
+%% the work being presented. Separate the keywords with commas.
+\begin{keywords}
+  LaTeX class \sep
+  paper template \sep
+  paper formatting \sep
+  CEUR-WS
+\end{keywords}
+
+%%
+%% This command processes the author and affiliation and title
+%% information and builds the first part of the formatted document.
 \maketitle
-    \include{sections/01-intro}
-    \include{figs/model-diagram}
+
+\section{Introduction}
+
+CEUR-WS's article template provides a consistent \LaTeX{} style for
+use across CEUR-WS publications, and incorporates accessibility and
+metadata-extraction functionality. This document will explain the
+major features of the document class.
+
+If you are new to publishing with CEUR-WS, this document is a valuable
+guide to the process of preparing your work for publication.
+
+The ``\verb|ceurart|'' document class can be used to prepare articles
+for any CEUR-WS publication, and for any stage of publication, from
+review to final ``camera-ready'' copy with {\itshape very} few changes
+to the source.
+
+This class depends on the following packages
+for its proper functioning:
+
+\begin{itemize}
+\item \verb|natbib.sty| for citation processing;
+\item \verb|geometry.sty| for margin settings;
+\item \verb|graphicx.sty| for graphics inclusion;
+\item \verb|hyperref.sty| optional package if hyperlinking is required in
+  the document;
+\item \verb|fontawesome5.sty| optional package for bells and whistles.
+\end{itemize}
+
+All the above packages are part of any
+standard \LaTeX{} installation.
+Therefore, the users need not be
+bothered about downloading any extra packages.
+
+\section{Modifications}
+
+Modifying the template --- including but not limited to: adjusting
+margins, typeface sizes, line spacing, paragraph and list definitions,
+and the use of the \mintinline{latex}{\vspace} command to manually adjust the
+vertical spacing between elements of your work --- is not allowed.
+
+\section{Template parameters}
+
+There are a number of template
+parameters which modify some part of the \verb|ceurart| document class.
+This parameters are enclosed in square
+brackets and are a part of the \mintinline{latex}{\documentclass} command:
+\begin{minted}{latex}
+  \documentclass[parameter]{ceurart}
+\end{minted}
+
+Frequently-used parameters, or combinations of parameters, include:
+\begin{itemize}
+\item \mintinline{latex}{twocolumn} : Two column layout.
+\item \mintinline{latex}{hf} : Enable header and footer\footnote{You can enable
+    the display of page numbers in the final version of the entire
+    collection. In this case, you should adhere to the end-to-end
+    pagination of individual papers.}.
+\end{itemize}
+
+\section{Front matter}
+
+\subsection{Title Information}
+
+The titles of papers should be either all use the emphasizing
+capitalized style or they should all use the regular English (or
+native language) style. It does not make a good impression if you or
+your authors mix the styles.
+
+Use the \mintinline{latex}{\title} command to define the title of your work. Do not
+insert line breaks in your title.
+
+\subsection{Title variants}
+
+\mintinline{latex}{\title} command have the below options:
+\begin{itemize}
+\item \verb|title|: Document title. This is default option. 
+\begin{minted}{latex}
+\title[mode=title]{This is a title}
+\end{minted}
+You can just omit it, like as follows:
+\begin{minted}{latex}
+\title{This is a title}
+\end{minted}
+
+\item \verb|alt|: Alternate title.
+\begin{minted}{latex}
+\title[mode=alt]{This is a alternate title}
+\end{minted}
+
+\item \verb|sub|: Sub title.
+\begin{minted}{latex}
+\title[mode=sub]{This is a sub title}
+\end{minted}
+You can just use \mintinline{latex}{\subtitle} command, as follows:
+\begin{minted}{latex}
+\subtitle{This is a sub title}
+\end{minted}
+
+\item \verb|trans|: Translated title.
+\begin{minted}{latex}
+\title[mode=trans]{This is a translated title}
+\end{minted}
+
+\item \verb|transsub|: Translated sub title.
+\begin{minted}{latex}
+\title[mode=transsub]{This is a translated sub title}
+\end{minted}
+\end{itemize}
+
+\subsection{Authors and Affiliations}
+
+Each author must be defined separately for accurate metadata
+identification. Multiple authors may share one affiliation. Authors'
+names should not be abbreviated; use full first names wherever
+possible. Include authors' e-mail addresses whenever possible.
+
+\mintinline{latex}{\author} command have the below options:
+
+\begin{itemize}
+\item \mintinline{latex}{style} : Style of author name (chinese)
+\item \mintinline{latex}{prefix} : Prefix
+\item \mintinline{latex}{suffix} : Suffix
+\item \mintinline{latex}{degree} : Degree
+\item \mintinline{latex}{role} : Role
+\item \mintinline{latex}{orcid} : ORCID
+\item \mintinline{latex}{email} : E-mail
+\item \mintinline{latex}{url} : URL
+\end{itemize}
+
+Author names can have some kinds of marks and notes:
+\begin{itemize}
+\item affiliation mark: \mintinline{latex}{\author[<num>]}.
+\end{itemize}
+
+The author names and affiliations could be formatted in two ways:
+\begin{enumerate}
+\item Group the authors per affiliation.
+\item Use an explicit mark to indicate the affiliations.
+\end{enumerate}
+
+Author block example:
+\begin{minted}{latex}
+\author[1,2]{Author Name}[%
+    prefix=Prof.,
+    degree=D.Sc.,
+    role=Researcher,
+    orcid=0000-0000-000-0000,
+    email=name@example.com,
+    url=https://name.example.com
+]
+
+\address[1]{Affiliation #1}
+\address[2]{Affiliation #2}
+\end{minted}
+
+\subsection{Abstract and Keywords}
+
+Abstract shall be entered in an environment that starts
+with \mintinline{latex}{\begin{abstract}} and ends with
+\mintinline{latex}{\end{abstract}}.
+
+\begin{minted}{latex}
+\begin{abstract}
+  This is an abstract.
+\end{abstract}
+\end{minted}
+
+The key words are enclosed in a \mintinline{latex}{keywords}
+environment. Use \mintinline{latex}{\sep} to separate keywords.
+
+\begin{minted}{latex}
+\begin{keywords}
+  First keyword \sep 
+  Second keyword \sep 
+  Third keyword \sep 
+  Fourth keyword
+\end{keywords}
+\end{minted}
+
+At the end of front matter add \mintinline{latex}{\maketitle} command.
+
+\subsection{Various Marks in the Front Matter}
+
+The front matter becomes complicated due to various kinds
+of notes and marks to the title and author names. Marks in
+the title will be denoted by a star ($\star$) mark;
+footnotes are denoted by super scripted Arabic numerals,
+corresponding author by an Conformal asterisk (*) mark.
+
+\subsubsection{Title marks}
+
+Title mark can be entered by the command, \mintinline{latex}{\tnotemark[<num>]}
+and the corresponding text can be entered with the command
+\mintinline{latex}{\tnotetext[<num>]{<text>}}. An example will be:
+
+\begin{minted}{latex}
+\title{A better way to format your document for CEUR-WS}
+
+\tnotemark[1]
+\tnotetext[1]{You can use this document as the template for preparing your
+  publication. We recommend using the latest version of the ceurart style.}
+\end{minted}
+
+\mintinline{latex}{\tnotemark} and \mintinline{latex}{\tnotetext} can be anywhere in
+the front matter, but should be before \mintinline{latex}{\maketitle} command.
+
+\subsubsection{Author marks}
+
+Author names can have some kinds of marks and notes:
+\begin{itemize}
+\item footnote mark : \mintinline{latex}{\fnmark[<num>]}
+\item footnote text : \mintinline{latex}{\fntext[<num>]{<text>}}
+\item corresponding author mark : \mintinline{latex}{\cormark[<num>]}
+\item corresponding author text : \mintinline{latex}{\cortext[<num>]{<text>}}
+\end{itemize}
+
+\subsubsection{Other marks}
+
+At times, authors want footnotes which leave no marks in
+the author names. The note text shall be listed as part of
+the front matter notes. Class files provides
+\mintinline{latex}{\nonumnote} for this purpose. The usage
+\begin{minted}{latex}
+\nonumnote{<text>}
+\end{minted}
+and should be entered anywhere before the \mintinline{latex}{\maketitle}
+command for this to take effect.
+
+\section{Sectioning Commands}
+
+Your work should use standard \LaTeX{} sectioning commands:
+\mintinline{latex}{\section}, \mintinline{latex}{\subsection},
+\mintinline{latex}{\subsubsection}, and
+\mintinline{latex}{\paragraph}. They should be numbered; do not remove
+the numbering from the commands.
+
+Simulating a sectioning command by setting the first word or words of
+a paragraph in boldface or italicized text is not allowed.
+
+\section{Tables}
+
+The ``\verb|ceurart|'' document class includes the ``\verb|booktabs|''
+package --- \url{https://ctan.org/pkg/booktabs} --- for preparing
+high-quality tables.
+
+Table captions are placed \textit{above} the table.
+
+Because tables cannot be split across pages, the best placement for
+them is typically the top of the page nearest their initial cite.  To
+ensure this proper ``floating'' placement of tables, use the
+environment \verb|table| to enclose the table's contents and the
+table caption. The contents of the table itself must go in the
+\verb|tabular| environment, to be aligned properly in rows and
+columns, with the desired horizontal and vertical rules.
+
+Immediately following this sentence is the point at which
+Table~\ref{tab:freq} is included in the input file; compare the
+placement of the table here with the table in the printed output of
+this document.
+
+\begin{table*}
+  \caption{Frequency of Special Characters}
+  \label{tab:freq}
+  \begin{tabular}{ccl}
+    \toprule
+    Non-English or Math&Frequency&Comments\\
+    \midrule
+    \O & 1 in 1,000& For Swedish names\\
+    $\pi$ & 1 in 5& Common in math\\
+    \$ & 4 in 5 & Used in business\\
+    $\Psi^2_1$ & 1 in 40,000& Unexplained usage\\
+  \bottomrule
+\end{tabular}
+\end{table*}
+
+To set a wider table, which takes up the whole width of the page's
+live area, use the environment \verb|table*| to enclose the table's
+contents and the table caption.  As with a single-column table, this
+wide table will ``float'' to a location deemed more
+desirable. Immediately following this sentence is the point at which
+Table~\ref{tab:commands} is included in the input file; again, it is
+instructive to compare the placement of the table here with the table
+in the printed output of this document.
+
+\begin{table}
+  \caption{Some Typical Commands}
+  \label{tab:commands}
+  \begin{tabular}{ccl}
+    \toprule
+    Command &A Number & Comments\\
+    \midrule
+    \texttt{{\char'134}author} & 100& Author \\
+    \texttt{{\char'134}table}& 300 & For tables\\
+    \texttt{{\char'134}table*}& 400& For wider tables\\
+    \bottomrule
+  \end{tabular}
+\end{table}
+
+\section{Math Equations}
+
+You may want to display math equations in three distinct styles:
+inline, numbered or non-numbered display.  Each of the three are
+discussed in the next sections.
+
+\subsection{Inline (In-text) Equations}
+
+A formula that appears in the running text is called an inline or
+in-text formula.  It is produced by the \verb|math| environment,
+which can be invoked with the usual
+\mintinline{latex}{\begin} \ldots \mintinline{latex}{\end} construction or with
+the short form \verb|$| \ldots \verb|$|. You can use any of the symbols
+and structures, from $\alpha$ to $\omega$, available in
+\LaTeX~\cite{Lamport:LaTeX};
+this section will simply show a few
+examples of in-text equations in context. Notice how this equation:
+\begin{math}
+  \lim_{n\rightarrow \infty} \frac{1}{n} = 0,
+\end{math}
+set here in in-line math style, looks slightly different when
+set in display style.  (See next section).
+
+\subsection{Display Equations}
+
+A numbered display equation---one set off by vertical space from the
+text and centered horizontally---is produced by the \verb|equation|
+environment. An unnumbered display equation is produced by the
+\verb|displaymath| environment.
+
+Again, in either environment, you can use any of the symbols and
+structures available in \LaTeX{}; this section will just give a couple
+of examples of display equations in context.  First, consider the
+equation, shown as an inline equation above:
+\begin{equation}
+  \lim_{n\rightarrow \infty} \frac{1}{n} = 0.
+\end{equation}
+Notice how it is formatted somewhat differently in
+the \verb|displaymath|
+environment.  Now, we'll enter an unnumbered equation:
+\begin{displaymath}
+  S_{n} = \sum_{i=1}^{n} x_{i} ,
+\end{displaymath}
+and follow it with another numbered equation:
+\begin{equation}
+  \lim_{x \to 0} (1 + x)^{1/x} = e
+\end{equation}
+just to demonstrate \LaTeX's able handling of numbering.
+
+\section{Figures}
+
+The ``\verb|figure|'' environment should be used for figures. One or
+more images can be placed within a figure. If your figure contains
+third-party material, you must clearly identify it as such, as shown
+in the example below.
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{sample-franklin}
+  \caption{1907 Franklin Model D roadster. Photograph by Harris \&
+    Ewing, Inc. [Public domain], via Wikimedia
+    Commons. (\url{https://goo.gl/VLCRBB}).}
+\end{figure}
+
+Your figures should contain a caption which describes the figure to
+the reader. Figure captions go below the figure. Your figures should
+also include a description suitable for screen readers, to
+assist the visually-challenged to better understand your work.
+
+Figure captions are placed below the figure.
+
+\section{Citations and Bibliographies}
+
+The use of Bib\TeX{} for the preparation and formatting of one's
+references is strongly recommended. Authors' names should be complete
+--- use full first names (``Donald E. Knuth'') not initials
+(``D. E. Knuth'') --- and the salient identifying features of a
+reference should be included: title, year, volume, number, pages,
+article DOI, etc.
+
+The bibliography is included in your source document with these two
+commands, placed just before the \mintinline{latex}{\end{document}}
+command:
+\begin{minted}{latex}
+\bibliography{bibfile}
+\end{minted}
+where ``\verb|bibfile|'' is the name, without the ``\verb|.bib|''
+suffix, of the Bib\TeX{} file.
+
+\subsection{Some examples}
+
+A paginated journal article \cite{Abril07}, an enumerated journal
+article \cite{Cohen07}, a reference to an entire issue
+\cite{JCohen96}, a monograph (whole book) \cite{Kosiur01}, a
+monograph/whole book in a series (see 2a in spec. document)
+\cite{Harel79}, a divisible-book such as an anthology or compilation
+\cite{Editor00} followed by the same example, however we only output
+the series if the volume number is given \cite{Editor00a} (so series
+should not be present since it has no vol. no.), a chapter in a
+divisible book \cite{Spector90}, a chapter in a divisible book in a
+series \cite{Douglass98}, a multi-volume work as book \cite{Knuth97},
+an article in a proceedings (of a conference, symposium, workshop for
+example) (paginated proceedings article) \cite{Andler79}, a
+proceedings article with all possible elements \cite{Smith10}, an
+example of an enumerated proceedings article \cite{VanGundy07}, an
+informally published work \cite{Harel78}, a doctoral dissertation
+\cite{Clarkson85}, a master's thesis: \cite{anisi03}, an online
+document / world wide web resource \cite{Thornburg01, Ablamowicz07,
+  Poker06}, a video game (Case 1) \cite{Obama08} and (Case 2)
+\cite{Novak03} and \cite{Lee05} and (Case 3) a patent
+\cite{JoeScientist001}, work accepted for publication \cite{rous08},
+prolific author \cite{SaeediMEJ10} and \cite{SaeediJETC10}. Other
+cites might contain `duplicate' DOI and URLs (some SIAM articles)
+\cite{Kirschmer:2010:AEI:1958016.1958018}. Multi-volume works as books
+\cite{MR781536} and \cite{MR781537}. A couple of citations with DOIs:
+\cite{2004:ITE:1009386.1010128,Kirschmer:2010:AEI:1958016.1958018}. Online
+citations: \cite{TUGInstmem, Thornburg01, R, UMassCitations}.
+
+\section{Acknowledgments}
+
+Identification of funding sources and other support, and thanks to
+individuals and groups that assisted in the research and the
+preparation of the work should be included in an acknowledgment
+section, which is placed just before the reference section in your
+document.
+
+This section has a special environment:
+\begin{minted}{latex}
+\begin{acknowledgments}
+  These are different acknowledgments.
+\end{acknowledgments}
+\end{minted}
+so that the information contained therein can be more easily collected
+during the article metadata extraction phase, and to ensure
+consistency in the spelling of the section heading.
+
+Authors should not prepare this section as a numbered or unnumbered
+\mintinline{latex}{\section}; please use the ``\mintinline{latex}{acknowledgments}'' environment.
+
+\section{Appendices}
+
+If your work needs an appendix, add it before the
+``\mintinline{latex}{\end{document}}'' command at the conclusion of your source
+document.
+
+Start the appendix with the ``\mintinline{latex}{\appendix}'' command:
+\begin{minted}{latex}
+\appendix
+\end{minted}
+and note that in the appendix, sections are lettered, not
+numbered.
+
+%%
+%% The acknowledgments section is defined using the "acknowledgments" environment
+%% (and NOT an unnumbered section). This ensures the proper
+%% identification of the section in the article metadata, and the
+%% consistent spelling of the heading.
+\begin{acknowledgments}
+  Thanks to the developers of ACM consolidated LaTeX styles
+  \url{https://github.com/borisveytsman/acmart} and to the developers
+  of Elsevier updated \LaTeX{} templates
+  \url{https://www.ctan.org/tex-archive/macros/latex/contrib/els-cas-templates}.  
+\end{acknowledgments}
+
+%%
+%% Define the bibliography file to be used
+\bibliography{sample-ceur}
+
+%%
+%% If your work has an appendix, this is the place to put it.
+\appendix
+
+\section{Online Resources}
+
+The sources for the ceur-art style are available via
+\begin{itemize}
+\item \href{https://github.com/yamadharma/ceurart}{GitHub},
+% \item \href{https://www.overleaf.com/project/5e76702c4acae70001d3bc87}{Overleaf},
+\item
+  \href{https://www.overleaf.com/latex/templates/template-for-submissions-to-ceur-workshop-proceedings-ceur-ws-dot-org/pkfscdkgkhcq}{Overleaf
+    template}.
+\end{itemize}
+
 \end{document}
+
+%%
+%% End of file

### 2024-05-25 paper/figs/model-diagram.tex
refactored the latex document into sections in subdirectories, added to diagram, updated some details

@@
 \tikzstyle{databox} = [rectangle, minimum width=40mm, minimum height=12mm, text centered, draw=red!70, fill=red!10, text=fontcolor!70]
-\tikzstyle{convnode} = [ellipse, minimum width=5mm, minimum height=5mm, fill=customteal!40, draw=customteal!80]
+\tikzstyle{convnode} = [rectangle, rounded corners=3mm, minimum width=6mm, minimum height=6mm, fill=customteal!40, draw=customteal!80]
+\tikzstyle{output} = [rectangle, minimum width=30mm, minimum height=10mm, fill=blue!20, draw=blue!50]
 \tikzstyle{sarrowxgb} = [->, thick, draw=customorange!70]
@@
 \tikzstyle{sarrowtr} = [->, thick, draw=customgreen!70]
+\tikzstyle{sarrowout} = [->, thick, draw=blue!50]
 
@@
 
+% Output
+\node (output) [output, right of=transformer, xshift=30mm, yshift=-15mm] {Output};
+\draw[sarrowout] (xgb.east) .. controls +(right:25mm) and +(north:50mm) .. (output.north);
+\draw[sarrowout] (conv.east) .. controls +(right:5mm) and +(north:20mm) .. (output.north);
+\draw[sarrowout] (transformer.east) .. controls +(right:15mm) and +(north:10mm) .. (output.north);
+
 \end{tikzpicture}

### 2024-05-25 paper/sections/00-title-abstract.tex
refactored the latex document into sections in subdirectories, added to diagram, updated some details

%%
%% Rights management information.
%% CC-BY is default license.
\copyrightyear{2024}
\copyrightclause{Copyright for this paper by its authors.
  Use permitted under Creative Commons License Attribution 4.0
  International (CC BY 4.0).}

%%
%% This command is for the conference information            %%%%%%% TODO !!! UPDATE THIS %%%%%%
\conference{Woodstock'22: Symposium on the irreproducible science,
  June 07--11, 2022, Woodstock, NY}

%%
%% The "title" command
\title{A better way to format your document for CEUR-WS}

% Authors
\author[1]{Darren Rawlings}[%
email=abc@def.ghi,
url=https://startung.github.io/,
]
\address[1]{University of Groningen, Broerstraat 5, 9712 CP Groningen, Netherlands}

\author[]{Tim Chopard}[%
email=timchopard@pm.me,
url=http://cloudberries.io,
]

\fntext[1]{These authors contributed equally.}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  A clear and well-documented \LaTeX{} document is presented as an
  article formatted for publication by CEUR-WS in a conference
  proceedings. Based on the ``ceurart'' document class, this article
  presents and explains many of the common variations, as well as many
  of the formatting elements an author may use in the preparation of
  the documentation of their work.
\end{abstract}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\begin{keywords}
  LaTeX class \sep
  paper template \sep
  paper formatting \sep
  CEUR-WS
\end{keywords}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

### 2024-05-25 paper/sections/01-intro.tex
refactored the latex document into sections in subdirectories, added to diagram, updated some details

\section{Introduction}

\begin{figure}
    \begin{center}
        \input{figs/model-diagram.tex}
        \caption{This diagram is just here temporarily and will be in a more relevant section when complete}
    \end{center}
\end{figure}

### 2024-05-25 paper/sections/99-template.tex
refactored the latex document into sections in subdirectories, added to diagram, updated some details

\section{Introduction}

CEUR-WS's article template provides a consistent \LaTeX{} style for
use across CEUR-WS publications, and incorporates accessibility and
metadata-extraction functionality. This document will explain the
major features of the document class.

If you are new to publishing with CEUR-WS, this document is a valuable
guide to the process of preparing your work for publication.

The ``\verb|ceurart|'' document class can be used to prepare articles
for any CEUR-WS publication, and for any stage of publication, from
review to final ``camera-ready'' copy with {\itshape very} few changes
to the source.

This class depends on the following packages
for its proper functioning:

\begin{itemize}
\item \verb|natbib.sty| for citation processing;
\item \verb|geometry.sty| for margin settings;
\item \verb|graphicx.sty| for graphics inclusion;
\item \verb|hyperref.sty| optional package if hyperlinking is required in
  the document;
\item \verb|fontawesome5.sty| optional package for bells and whistles.
\end{itemize}

All the above packages are part of any
standard \LaTeX{} installation.
Therefore, the users need not be
bothered about downloading any extra packages.

\section{Modifications}

Modifying the template --- including but not limited to: adjusting
margins, typeface sizes, line spacing, paragraph and list definitions,
and the use of the \mintinline{latex}{\vspace} command to manually adjust the
vertical spacing between elements of your work --- is not allowed.

\section{Template parameters}

There are a number of template
parameters which modify some part of the \verb|ceurart| document class.
This parameters are enclosed in square
brackets and are a part of the \mintinline{latex}{\documentclass} command:
\begin{minted}{latex}
  \documentclass[parameter]{ceurart}
\end{minted}

Frequently-used parameters, or combinations of parameters, include:
\begin{itemize}
\item \mintinline{latex}{twocolumn} : Two column layout.
\item \mintinline{latex}{hf} : Enable header and footer\footnote{You can enable
    the display of page numbers in the final version of the entire
    collection. In this case, you should adhere to the end-to-end
    pagination of individual papers.}.
\end{itemize}

\section{Front matter}

\subsection{Title Information}

The titles of papers should be either all use the emphasizing
capitalized style or they should all use the regular English (or
native language) style. It does not make a good impression if you or
your authors mix the styles.

Use the \mintinline{latex}{\title} command to define the title of your work. Do not
insert line breaks in your title.

\subsection{Title variants}

\mintinline{latex}{\title} command have the below options:
\begin{itemize}
\item \verb|title|: Document title. This is default option. 
\begin{minted}{latex}
\title[mode=title]{This is a title}
\end{minted}
You can just omit it, like as follows:
\begin{minted}{latex}
\title{This is a title}
\end{minted}

\item \verb|alt|: Alternate title.
\begin{minted}{latex}
\title[mode=alt]{This is a alternate title}
\end{minted}

\item \verb|sub|: Sub title.
\begin{minted}{latex}
\title[mode=sub]{This is a sub title}
\end{minted}
You can just use \mintinline{latex}{\subtitle} command, as follows:
\begin{minted}{latex}
\subtitle{This is a sub title}
\end{minted}

\item \verb|trans|: Translated title.
\begin{minted}{latex}
\title[mode=trans]{This is a translated title}
\end{minted}

\item \verb|transsub|: Translated sub title.
\begin{minted}{latex}
\title[mode=transsub]{This is a translated sub title}
\end{minted}
\end{itemize}

\subsection{Authors and Affiliations}

Each author must be defined separately for accurate metadata
identification. Multiple authors may share one affiliation. Authors'
names should not be abbreviated; use full first names wherever
possible. Include authors' e-mail addresses whenever possible.

\mintinline{latex}{\author} command have the below options:

\begin{itemize}
\item \mintinline{latex}{style} : Style of author name (chinese)
\item \mintinline{latex}{prefix} : Prefix
\item \mintinline{latex}{suffix} : Suffix
\item \mintinline{latex}{degree} : Degree
\item \mintinline{latex}{role} : Role
\item \mintinline{latex}{orcid} : ORCID
\item \mintinline{latex}{email} : E-mail
\item \mintinline{latex}{url} : URL
\end{itemize}

Author names can have some kinds of marks and notes:
\begin{itemize}
\item affiliation mark: \mintinline{latex}{\author[<num>]}.
\end{itemize}

The author names and affiliations could be formatted in two ways:
\begin{enumerate}
\item Group the authors per affiliation.
\item Use an explicit mark to indicate the affiliations.
\end{enumerate}

Author block example:
\begin{minted}{latex}
\author[1,2]{Author Name}[%
    prefix=Prof.,
    degree=D.Sc.,
    role=Researcher,
    orcid=0000-0000-000-0000,
    email=name@example.com,
    url=https://name.example.com
]

\address[1]{Affiliation #1}
\address[2]{Affiliation #2}
\end{minted}

\subsection{Abstract and Keywords}

Abstract shall be entered in an environment that starts
with \mintinline{latex}{\begin{abstract}} and ends with
\mintinline{latex}{\end{abstract}}.

\begin{minted}{latex}
\begin{abstract}
  This is an abstract.
\end{abstract}
\end{minted}

The key words are enclosed in a \mintinline{latex}{keywords}
environment. Use \mintinline{latex}{\sep} to separate keywords.

\begin{minted}{latex}
\begin{keywords}
  First keyword \sep 
  Second keyword \sep 
  Third keyword \sep 
  Fourth keyword
\end{keywords}
\end{minted}

At the end of front matter add \mintinline{latex}{\maketitle} command.

\subsection{Various Marks in the Front Matter}

The front matter becomes complicated due to various kinds
of notes and marks to the title and author names. Marks in
the title will be denoted by a star ($\star$) mark;
footnotes are denoted by super scripted Arabic numerals,
corresponding author by an Conformal asterisk (*) mark.

\subsubsection{Title marks}

Title mark can be entered by the command, \mintinline{latex}{\tnotemark[<num>]}
and the corresponding text can be entered with the command
\mintinline{latex}{\tnotetext[<num>]{<text>}}. An example will be:

\begin{minted}{latex}
\title{A better way to format your document for CEUR-WS}

\tnotemark[1]
\tnotetext[1]{You can use this document as the template for preparing your
  publication. We recommend using the latest version of the ceurart style.}
\end{minted}

\mintinline{latex}{\tnotemark} and \mintinline{latex}{\tnotetext} can be anywhere in
the front matter, but should be before \mintinline{latex}{\maketitle} command.

\subsubsection{Author marks}

Author names can have some kinds of marks and notes:
\begin{itemize}
\item footnote mark : \mintinline{latex}{\fnmark[<num>]}
\item footnote text : \mintinline{latex}{\fntext[<num>]{<text>}}
\item corresponding author mark : \mintinline{latex}{\cormark[<num>]}
\item corresponding author text : \mintinline{latex}{\cortext[<num>]{<text>}}
\end{itemize}

\subsubsection{Other marks}

At times, authors want footnotes which leave no marks in
the author names. The note text shall be listed as part of
the front matter notes. Class files provides
\mintinline{latex}{\nonumnote} for this purpose. The usage
\begin{minted}{latex}
\nonumnote{<text>}
\end{minted}
and should be entered anywhere before the \mintinline{latex}{\maketitle}
command for this to take effect.

\section{Sectioning Commands}

Your work should use standard \LaTeX{} sectioning commands:
\mintinline{latex}{\section}, \mintinline{latex}{\subsection},
\mintinline{latex}{\subsubsection}, and
\mintinline{latex}{\paragraph}. They should be numbered; do not remove
the numbering from the commands.

Simulating a sectioning command by setting the first word or words of
a paragraph in boldface or italicized text is not allowed.

\section{Tables}

The ``\verb|ceurart|'' document class includes the ``\verb|booktabs|''
package --- \url{https://ctan.org/pkg/booktabs} --- for preparing
high-quality tables.

Table captions are placed \textit{above} the table.

Because tables cannot be split across pages, the best placement for
them is typically the top of the page nearest their initial cite.  To
ensure this proper ``floating'' placement of tables, use the
environment \verb|table| to enclose the table's contents and the
table caption. The contents of the table itself must go in the
\verb|tabular| environment, to be aligned properly in rows and
columns, with the desired horizontal and vertical rules.

Immediately following this sentence is the point at which
Table~\ref{tab:freq} is included in the input file; compare the
placement of the table here with the table in the printed output of
this document.

\begin{table*}
  \caption{Frequency of Special Characters}
  \label{tab:freq}
  \begin{tabular}{ccl}
    \toprule
    Non-English or Math&Frequency&Comments\\
    \midrule
    \O & 1 in 1,000& For Swedish names\\
    $\pi$ & 1 in 5& Common in math\\
    \$ & 4 in 5 & Used in business\\
    $\Psi^2_1$ & 1 in 40,000& Unexplained usage\\
  \bottomrule
\end{tabular}
\end{table*}

To set a wider table, which takes up the whole width of the page's
live area, use the environment \verb|table*| to enclose the table's
contents and the table caption.  As with a single-column table, this
wide table will ``float'' to a location deemed more
desirable. Immediately following this sentence is the point at which
Table~\ref{tab:commands} is included in the input file; again, it is
instructive to compare the placement of the table here with the table
in the printed output of this document.

\begin{table}
  \caption{Some Typical Commands}
  \label{tab:commands}
  \begin{tabular}{ccl}
    \toprule
    Command &A Number & Comments\\
    \midrule
    \texttt{{\char'134}author} & 100& Author \\
    \texttt{{\char'134}table}& 300 & For tables\\
    \texttt{{\char'134}table*}& 400& For wider tables\\
    \bottomrule
  \end{tabular}
\end{table}

\section{Math Equations}

You may want to display math equations in three distinct styles:
inline, numbered or non-numbered display.  Each of the three are
discussed in the next sections.

\subsection{Inline (In-text) Equations}

A formula that appears in the running text is called an inline or
in-text formula.  It is produced by the \verb|math| environment,
which can be invoked with the usual
\mintinline{latex}{\begin} \ldots \mintinline{latex}{\end} construction or with
the short form \verb|$| \ldots \verb|$|. You can use any of the symbols
and structures, from $\alpha$ to $\omega$, available in
\LaTeX~\cite{Lamport:LaTeX};
this section will simply show a few
examples of in-text equations in context. Notice how this equation:
\begin{math}
  \lim_{n\rightarrow \infty} \frac{1}{n} = 0,
\end{math}
set here in in-line math style, looks slightly different when
set in display style.  (See next section).

\subsection{Display Equations}

A numbered display equation---one set off by vertical space from the
text and centered horizontally---is produced by the \verb|equation|
environment. An unnumbered display equation is produced by the
\verb|displaymath| environment.

Again, in either environment, you can use any of the symbols and
structures available in \LaTeX{}; this section will just give a couple
of examples of display equations in context.  First, consider the
equation, shown as an inline equation above:
\begin{equation}
  \lim_{n\rightarrow \infty} \frac{1}{n} = 0.
\end{equation}
Notice how it is formatted somewhat differently in
the \verb|displaymath|
environment.  Now, we'll enter an unnumbered equation:
\begin{displaymath}
  S_{n} = \sum_{i=1}^{n} x_{i} ,
\end{displaymath}
and follow it with another numbered equation:
\begin{equation}
  \lim_{x \to 0} (1 + x)^{1/x} = e
\end{equation}
just to demonstrate \LaTeX's able handling of numbering.

\section{Figures}

The ``\verb|figure|'' environment should be used for figures. One or
more images can be placed within a figure. If your figure contains
third-party material, you must clearly identify it as such, as shown
in the example below.
\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{sample-franklin}
  \caption{1907 Franklin Model D roadster. Photograph by Harris \&
    Ewing, Inc. [Public domain], via Wikimedia
    Commons. (\url{https://goo.gl/VLCRBB}).}
\end{figure}

Your figures should contain a caption which describes the figure to
the reader. Figure captions go below the figure. Your figures should
also include a description suitable for screen readers, to
assist the visually-challenged to better understand your work.

Figure captions are placed below the figure.

\section{Citations and Bibliographies}

The use of Bib\TeX{} for the preparation and formatting of one's
references is strongly recommended. Authors' names should be complete
--- use full first names (``Donald E. Knuth'') not initials
(``D. E. Knuth'') --- and the salient identifying features of a
reference should be included: title, year, volume, number, pages,
article DOI, etc.

The bibliography is included in your source document with these two
commands, placed just before the \mintinline{latex}{\end{document}}
command:
\begin{minted}{latex}
\bibliography{bibfile}
\end{minted}
where ``\verb|bibfile|'' is the name, without the ``\verb|.bib|''
suffix, of the Bib\TeX{} file.

\subsection{Some examples}

A paginated journal article \cite{Abril07}, an enumerated journal
article \cite{Cohen07}, a reference to an entire issue
\cite{JCohen96}, a monograph (whole book) \cite{Kosiur01}, a
monograph/whole book in a series (see 2a in spec. document)
\cite{Harel79}, a divisible-book such as an anthology or compilation
\cite{Editor00} followed by the same example, however we only output
the series if the volume number is given \cite{Editor00a} (so series
should not be present since it has no vol. no.), a chapter in a
divisible book \cite{Spector90}, a chapter in a divisible book in a
series \cite{Douglass98}, a multi-volume work as book \cite{Knuth97},
an article in a proceedings (of a conference, symposium, workshop for
example) (paginated proceedings article) \cite{Andler79}, a
proceedings article with all possible elements \cite{Smith10}, an
example of an enumerated proceedings article \cite{VanGundy07}, an
informally published work \cite{Harel78}, a doctoral dissertation
\cite{Clarkson85}, a master's thesis: \cite{anisi03}, an online
document / world wide web resource \cite{Thornburg01, Ablamowicz07,
  Poker06}, a video game (Case 1) \cite{Obama08} and (Case 2)
\cite{Novak03} and \cite{Lee05} and (Case 3) a patent
\cite{JoeScientist001}, work accepted for publication \cite{rous08},
prolific author \cite{SaeediMEJ10} and \cite{SaeediJETC10}. Other
cites might contain `duplicate' DOI and URLs (some SIAM articles)
\cite{Kirschmer:2010:AEI:1958016.1958018}. Multi-volume works as books
\cite{MR781536} and \cite{MR781537}. A couple of citations with DOIs:
\cite{2004:ITE:1009386.1010128,Kirschmer:2010:AEI:1958016.1958018}. Online
citations: \cite{TUGInstmem, Thornburg01, R, UMassCitations}.

\section{Acknowledgments}

Identification of funding sources and other support, and thanks to
individuals and groups that assisted in the research and the
preparation of the work should be included in an acknowledgment
section, which is placed just before the reference section in your
document.

This section has a special environment:
\begin{minted}{latex}
\begin{acknowledgments}
  These are different acknowledgments.
\end{acknowledgments}
\end{minted}
so that the information contained therein can be more easily collected
during the article metadata extraction phase, and to ensure
consistency in the spelling of the section heading.

Authors should not prepare this section as a numbered or unnumbered
\mintinline{latex}{\section}; please use the ``\mintinline{latex}{acknowledgments}'' environment.

\section{Appendices}

If your work needs an appendix, add it before the
``\mintinline{latex}{\end{document}}'' command at the conclusion of your source
document.

Start the appendix with the ``\mintinline{latex}{\appendix}'' command:
\begin{minted}{latex}
\appendix
\end{minted}
and note that in the appendix, sections are lettered, not
numbered.

%%
%% The acknowledgments section is defined using the "acknowledgments" environment
%% (and NOT an unnumbered section). This ensures the proper
%% identification of the section in the article metadata, and the
%% consistent spelling of the heading.
\begin{acknowledgments}
  Thanks to the developers of ACM consolidated LaTeX styles
  \url{https://github.com/borisveytsman/acmart} and to the developers
  of Elsevier updated \LaTeX{} templates
  \url{https://www.ctan.org/tex-archive/macros/latex/contrib/els-cas-templates}.  
\end{acknowledgments}

%%
%% Define the bibliography file to be used
\bibliography{sample-ceur}

%%
%% If your work has an appendix, this is the place to put it.
\appendix

\section{Online Resources}

The sources for the ceur-art style are available via
\begin{itemize}
\item \href{https://github.com/yamadharma/ceurart}{GitHub},
% \item \href{https://www.overleaf.com/project/5e76702c4acae70001d3bc87}{Overleaf},
\item
  \href{https://www.overleaf.com/latex/templates/template-for-submissions-to-ceur-workshop-proceedings-ceur-ws-dot-org/pkfscdkgkhcq}{Overleaf
    template}.
\end{itemize}

### 2024-05-28 paper/figs/model-diagram.tex
updates to diagram for easier editing
initialized method section

@@
 % Tikz shape definitions
-\tikzstyle{nnbox} = [rectangle, rounded corners=0.5mm, minimum width=35mm, minimum height=10mm, text centered, draw=customteal, fill=customteal!30, text=fontcolor]
-\tikzstyle{transformerbox} = [rectangle, rounded corners=0.5mm, minimum width=35mm, minimum height=10mm, text centered, draw=customgreen, fill=customgreen!30, text=fontcolor]
-\tikzstyle{xgbbox} = [rectangle, rounded corners=0.5mm, minimum width=35mm, minimum height=10mm, text centered, draw=customorange, fill=customorange!30, text=fontcolor]
-\tikzstyle{pcabox} = [rectangle, rounded corners=2mm, minimum width=20mm, minimum height=7mm, text centered, draw=customorange!70, fill=customorange!10, text=fontcolor!70]
-\tikzstyle{databox} = [rectangle, minimum width=40mm, minimum height=12mm, text centered, draw=red!70, fill=red!10, text=fontcolor!70]
+\tikzstyle{parentbox} = [rectangle, rounded corners=1mm] % TODO
+\tikzstyle{nnbox} = [parentbox, minimum width=35mm, minimum height=10mm, text centered, draw=customteal, fill=customteal!30, text=fontcolor]
+\tikzstyle{transformerbox} = [parentbox, minimum width=35mm, minimum height=10mm, text centered, draw=customgreen, fill=customgreen!30, text=fontcolor]
+\tikzstyle{xgbbox} = [parentbox, minimum width=35mm, minimum height=10mm, text centered, draw=customorange, fill=customorange!30, text=fontcolor]
+\tikzstyle{pcabox} = [parentbox, minimum width=20mm, minimum height=7mm, text centered, draw=customorange!70, fill=customorange!10, text=fontcolor!70]
+\tikzstyle{databox} = [parentbox, minimum width=40mm, minimum height=12mm, text centered, draw=red!70, fill=red!10, text=fontcolor!70]
 \tikzstyle{convnode} = [rectangle, rounded corners=3mm, minimum width=6mm, minimum height=6mm, fill=customteal!40, draw=customteal!80]
-\tikzstyle{output} = [rectangle, minimum width=30mm, minimum height=10mm, fill=blue!20, draw=blue!50]
-\tikzstyle{sarrowxgb} = [->, thick, draw=customorange!70]
-\tikzstyle{sarrownn} = [->, thick, draw=customteal!70]
-\tikzstyle{sarrowtr} = [->, thick, draw=customgreen!70]
-\tikzstyle{sarrowout} = [->, thick, draw=blue!50]
+\tikzstyle{output} = [parentbox, minimum width=30mm, minimum height=10mm, fill=blue!20, draw=blue!50]
+
+% Arrows
+\tikzstyle{parentarrow} = [-stealth, line width=0.5mm]
+\tikzstyle{sarrowxgb} = [parentarrow, draw=customorange!70]
+\tikzstyle{sarrownn} = [parentarrow, draw=customteal!70]
+\tikzstyle{sarrowtr} = [parentarrow, draw=customgreen!70]
+\tikzstyle{sarrowout} = [parentarrow, draw=blue!50]
 

### 2024-05-28 paper/sections/01-intro.tex
Add introduction first draft

@@
 \section{Introduction}
+
+In this research study, we aim to develop a model for predicting plant species in a specific location and time using various environmental factors as predictors.
+These predictors include satellite images, climatic time series, and other rasterized environmental data such as land cover, human footprint, bioclimatic variables, and soil characteristics.
+Our motivation behind this challenge is the potential usefulness of accurate plant species prediction in various scenarios related to biodiversity management and conservation, species identification and inventory tools, and education.
+
+We utilized a large-scale training dataset of approximately 5 million plant occurrences in Europe, as well as validation and test sets with over 5,000 and 20,000 plots, respectively.
+The predicted output will be multi-label, presence-absence data for all present species at each plot.
+The data covered over 10,000 different plant species, which created significant challendges associated with this task, including learning from single positive labels, dealing with strong class imbalance, multi-modal learning, and handling large-scale datasets.
+
+The potential applications of accurate plant species prediction are numerous.
+High-resolution maps of species composition and related biodiversity indicators can be created to aid in scientific ecology studies and conservation efforts.
+The accuracy of species identification tools can be improved by reducing the list of candidate species observable at a given site.
+Additionally, location-based recommendation services and educational applications with features such as quests or contextualized educational pathways can be developed to facilitate biodiversity inventories and promote environmental education.
+We believe that our research will contribute to the advancement of plant species prediction and its practical applications in various fields.
 

### 2024-05-28 paper/sections/04-method.tex
initial methods structure written

\clearpage  % TODO comment this out when finished

\section{Method}

In order to manage the multilabel classification required for this research we implemented an ensemble approach. 
This section details the individual architectures and the methods used to combine them.

\subsection{Architectures}

The core architectures used in this project were 18 layer ResNet, XGBoost and [TRANSFORMER]. The outputs of these were 
then weighted and combined before the maximum arguments were selected.

\subsubsection{XGBoost}
XGBoost is an open source gradient tree boosting package \cite{xgboost}. For this research we used the xgbregression model It has shown broad success on a number of 
\subsubsection{[Resnet]}
\subsubsection{[Transformer]}

\subsection{Process}
\subsubsection{Species scoring}
\subsubsection{Species counts}
\subsection{Metrics}

### 2024-05-28 paper/sections/03-data.tex
Add first draft of data

\section{Data}

Dataset Description

The training data comprises species observations and environmental data. Below, we explain the data in detail.

The species related training data comprises:

Presence-Absence (PA) surveys: including around 90 thousand surveys with roughly 10,000 species of the European flora. The presence-absence data (PA) is provided to compensate for the problem of false-absences of PO data and calibrate models to avoid associated biases.
Presence-Only (PO) occurrences: combines around five million observations from numerous datasets gathered from the Global Biodiversity Information Facility (GBIF, www.gbif.org). This data constitutes the larger piece of the training data and covers all countries of our study area, but it has been sampled opportunistically (without standardized sampling protocol), leading to various sampling biases. The local absence of a species among PO data doesn't mean it is truly absent. An observer might not have reported it because it was difficult to "see" it at this time of the year, to identify it as not a monitoring target, or just unattractive.

Environmental data

Besides species data, we provide spatialized geographic and environmental data as additional input variables (see Figure 1). More precisely, For each species observation location, we provide:

Satellite image patches: 3-band (RGB) and 1-band (NIR) 128x128 JPEG images, a color JPEG file for RGB data and a grayscale one for Near-Infrared images at 10m resolution. The source for these images is Sentinel2 remote sensing data pre-processed by the Ecodatacube platform.

Satellite time series: Up to 20 years of values for six satellite bands (R, G, B, NIR, SWIR1, and SWIR2). Each observation is associated with the time series of the satellite median point values over each season since the winter of 1999 for six satellite bands (R, G, B, NIR, SWIR1, and SWIR2). This data carries a high-resolution local signature of the past 20 years' succession of seasonal vegetation changes, potential extreme natural events (fires), or land use changes. The original satellite data has a resolution of 30m per pixel. The source for this is the Landsat remote sensing data pre-processed by the Ecodatacube platform

Environmental rasters Various climatic, pedologic, land use, and human footprint variables at the European scale. We provide scalar values, time-series, and original rasters from which you may extract local 2D images.

Four climatic variables computed monthly (mean, minimum and maximum temperature, and total precipitation) from January 2000 to December 2019, yielding 960 low-resolution (30 arcsec ~ 1 kilometer) rasters covering Europe. The source for these rasters is the CHELSA climate dataset.

Environmental rasters, for each observation, we were provided additional environmental data such as GeoTIFF rasters and scalar values already extracted from the rasters. We provide CSV files, one per band raster type, i.e., Climate, Elevation, Human Footprint, LandCover, and SoilGrids.

Bioclimatic rasters: 19 low-resolution rasters covering Europe; commonly used in species distribution modeling. Provided in longitude/latitude coordinates (WGS84). These were provided as GeoTIFF files with compression and CSV file with extracted values, with a resolution of 30 arcsec (~ 1 kilometer). The source for these rasters is the CHELSA climate dataset.

Soil rasters: Nine pedologic low-resolution rasters covering Europe. Provided variables describe the soil properties from 5 to 15cm depth and are determinant of plant species distributions. Check the definition.txt file about the provided variables (e.g., pH, clay, organic carbon and nitrogen contents, etc.). The format is GeoTIFF files with compression and CSV file with extracted values, with a resolution of ~1 kilometer. The source for these rasters is Soilgrids.

Elevation: High-resolution raster covering Europe. Provided as a GeoTIFF file and CSV file with extracted values, with a resolution of 1 arc second (~30 meters). The source for this raster is the ASTER Global Digital Elevation Model V3.

Land Cover: A medium-resolution multi-band land cover raster covering Europe. Each band describes either the land cover class prediction or its confidence under various classifications. We recommend the use of IGBP (17 classes) or LCCS (43 classes) layers, often used in species distribution modeling. The format is GeoTIFF file with compression and CSV file with extracted values, with a resolution of ~500 meters. The source for this raster is MODIS Terra+Aqua 500m.

Human footprint: Several low-resolution rasters describing human footprint, encapsulating seven pressures on the environment (e.g., nighlight level, population density) induced by human presence and activity, are provided for two time periods, the early 90's (~1993) and late 2000' (~2009). We provide two summary rasters combining all human pressures and two detailed rasters per pressure, which avoid an arbitrary degradation of the original data. The format is GeoTIFF files with compression and CSV file with extracted values, with a resolution of ~1 kilometer. The source for these rasters is \cite{human-footprint-dataset-2016}.

### 2024-05-28 paper/figs/count-diagram.tex
updates to method

% Tikz shape definitions
\tikzstyle{parentbox} = [rectangle, rounded corners=1mm] % TODO
\tikzstyle{nnbox} = [parentbox, minimum width=35mm, minimum height=10mm, text centered, draw=customteal, fill=customteal!30, text=fontcolor]
\tikzstyle{transformerbox} = [parentbox, minimum width=35mm, minimum height=10mm, text centered, draw=customgreen, fill=customgreen!30, text=fontcolor]
\tikzstyle{xgbbox} = [parentbox, minimum width=35mm, minimum height=10mm, text centered, draw=customorange, fill=customorange!30, text=fontcolor]
\tikzstyle{pcabox} = [parentbox, minimum width=20mm, minimum height=7mm, text centered, draw=customorange!70, fill=customorange!10, text=fontcolor!70]
\tikzstyle{databox} = [parentbox, minimum width=40mm, minimum height=12mm, text centered, draw=red!70, fill=red!10, text=fontcolor!70]
\tikzstyle{convnode} = [rectangle, rounded corners=3mm, minimum width=6mm, minimum height=6mm, fill=customteal!40, draw=customteal!80]
\tikzstyle{output} = [parentbox, minimum width=30mm, minimum height=10mm, fill=blue!20, draw=blue!50]

\tikzstyle{baseens} = [parentbox, minimum height=30mm, minimum height=20mm, fill=customorange!30, draw=customorange!70]

% Arrows
\tikzstyle{parentarrow} = [-stealth, line width=0.5mm]
\tikzstyle{sarrowxgb} = [parentarrow, draw=customorange!70]
\tikzstyle{sarrownn} = [parentarrow, draw=customteal!70]
\tikzstyle{sarrowtr} = [parentarrow, draw=customgreen!70]
\tikzstyle{sarrowout} = [parentarrow, draw=blue!50]

\begin{tikzpicture}

% Data
\node (bioclim) [databox] {Bioclimatic Rasters};
\node (landsat) [databox, below of=bioclim, yshift=-7mm] {LandSat};
\node (metadata) [databox, below of=landsat, yshift=-7mm] {Metadata};
\node (sentinel) [databox, below of=metadata, yshift=-7mm] {Sentinel};

\node (baseens) [baseens, right of=bioclim, yshift=-43mm, xshift=40mm] {Score Ensemble};
\draw[sarrowxgb] (bioclim.east) .. controls +(right:5mm) and +(left:8mm) .. (baseens.west);
\draw[sarrowxgb] (landsat.east) .. controls +(right:5mm) and +(left:8mm) .. (baseens.west);
\draw[sarrowxgb] (metadata.east) .. controls +(right:5mm) and +(left:8mm) .. (baseens.west);
\draw[sarrowxgb] (sentinel.east) .. controls +(right:5mm) and +(left:8mm) .. (baseens.west);

    % % PCAs
% \node (pca0) [pcabox, right of=bioclim, xshift=30mm, yshift=30mm] {PCA$_0$}; 
% \node (pca1) [pcabox, below of=pca0] {PCA$_1$}; 
% \node (vertellipsis0) [below of=pca1, yshift=3mm] {\Large{\textcolor{fontcolor!70}{$\cdot$}}};
% \node (vertellipsis1) [below of=vertellipsis0, yshift=9mm] {\Large{\textcolor{fontcolor!70}{$\cdot$}}};
% \node (vertellipsis2) [below of=vertellipsis1, yshift=9mm] {\Large{\textcolor{fontcolor!70}{$\cdot$}}};
% \node (pca2) [pcabox, below of=vertellipsis2, yshift=3mm] {PCA$_n$}; 
% \draw[sarrowxgb] (bioclim.east) .. controls +(right:5mm) and +(left:8mm) .. (pca0.west);
% \draw[sarrowxgb] (bioclim.east) .. controls +(right:5mm) and +(left:8mm) .. (pca1.west);
% \draw[sarrowxgb] (bioclim.east) .. controls +(right:5mm) and +(left:8mm) .. (pca2.west);

% % XGB
% \node (xgb) [xgbbox, right of=pca0, xshift=30mm, yshift=-15mm] {XGBoostRegressor};
% \draw[sarrowxgb] (pca0.east) .. controls +(right:5mm) and +(left:8mm) .. (xgb.west);
% \draw[sarrowxgb] (pca1.east) .. controls +(right:5mm) and +(left:8mm) .. (xgb.west);
% \draw[sarrowxgb] (pca2.east) .. controls +(right:5mm) and +(left:8mm) .. (xgb.west);
% \draw[sarrowxgb] (landsat.east) .. controls +(right:35mm) and +(left:10mm) .. (xgb.west);
% \draw[sarrowxgb] (metadata.east) .. controls +(right:35mm) and +(left:10mm) .. (xgb.west);

% % ResNet
% \node (nn0) [nnbox, below of=xgb, yshift=-15mm] {ResNet18};
% \node (nn1) [nnbox, below of=nn0, yshift=-3mm] {ResNet18};
% \node (nn2) [nnbox, below of=nn1, yshift=-3mm] {ResNet18};
% \node (conv) [convnode, right of=nn1, xshift=20mm] {};
% \draw[sarrownn] (bioclim.east) .. controls +(right:25mm) and +(left:20mm) .. (nn0.west);
% \draw[sarrownn] (landsat.east) .. controls +(right:25mm) and +(left:20mm) .. (nn1.west);
% \draw[sarrownn] (sentinel.east) .. controls +(right:25mm) and +(left:20mm) .. (nn2.west);
% \draw[sarrownn] (nn0.east) .. controls +(right:7mm) and +(left:7mm) .. (conv.west);
% \draw[sarrownn] (nn1.east) .. controls +(right:7mm) and +(left:7mm) .. (conv.west);
% \draw[sarrownn] (nn2.east) .. controls +(right:7mm) and +(left:7mm) .. (conv.west);

% % Transformer
% \node (transformer) [transformerbox, below of=nn2, yshift=-15mm] {Transformer};
% \draw[sarrowtr] (bioclim.east) .. controls +(right:25mm) and +(left:20mm) .. (transformer.west);
% \draw[sarrowtr] (landsat.east) .. controls +(right:25mm) and +(left:20mm) .. (transformer.west);
% \draw[sarrowtr] (metadata.east) .. controls +(right:25mm) and +(left:20mm) .. (transformer.west);
% \draw[sarrowtr] (sentinel.east) .. controls +(right:25mm) and +(left:20mm) .. (transformer.west);

% % Output
% \node (output) [output, right of=transformer, xshift=30mm, yshift=-15mm] {Output};
% \draw[sarrowout] (xgb.east) .. controls +(right:25mm) and +(north:50mm) .. (output.north);
% \draw[sarrowout] (conv.east) .. controls +(right:5mm) and +(north:20mm) .. (output.north);
% \draw[sarrowout] (transformer.east) .. controls +(right:15mm) and +(north:10mm) .. (output.north);

\end{tikzpicture}

### 2024-05-28 paper/sections/04-method.tex
updates to method

@@
 
+\subsubsection{Principal Component Analysis}
+Principal Component Analysis (PCA) is a method for reducing dimensionality
+
 \subsubsection{XGBoost}
-XGBoost is an open source gradient tree boosting package \cite{xgboost}. For this research we used the xgbregression model It has shown broad success on a number of 
-\subsubsection{[Resnet]}
+XGBoost is an open source gradient tree boosting package~\cite{xgboost}. 
+For this research we used the xgbregression model It has shown broad success accross a range of tasks, performing on par with or better than most equivalent and Automated Machine learning approaches~\cite{xgbcomp}.
+\subsubsection{ResNet}
+ResNet (Residual Network) is an architecture for deep neural networks that speeds up the training process through the use of residual connections~\cite{resnet}. 
+For the purposes of this research we focused on ResNet18, the 18 layer deep variation with some modifications to accomodate the shape of the input data and the required output. 
 \subsubsection{[Transformer]}
@@
 \subsection{Process}
+
+\newcommand{\nspecies}{N_{\text{species}}}
+
+There were two key steps to the process of species selection. First the core models which generated pseudo-probabilities and were then weighted and combined. 
+Second separeate models were used to determine the expected number $(\nspecies)$ of species per survey. 
+These steps came together with the expected counts being used to select the top  $\nspecies$ psuedo-probabilities per survey.
+
 \subsubsection{Species scoring}
+\begin{figure}
+    \begin{center}
+        \input{figs/model-diagram.tex}
+        \caption{The structure of the scoring ensemble from data through to individual species score outputs.}
+    \end{center}\label{fig:score-structure}
+\end{figure}
 \subsubsection{Species counts}
+\begin{figure}
+    \begin{center}
+        \input{figs/count-diagram.tex}
+        \caption{The counts ensemble through to final outputs.}
+    \end{center}\label{fig:count-structure}
+\end{figure}
 \subsection{Metrics}
+
+\newcommand{\fp}{\text{FP}}
+\newcommand{\fn}{\text{FN}}
+\newcommand{\tp}{\text{FP}}
+
+The main scoring metric used was the micro-averaged F1 score as shown in equation \ref{eq:mf1} which is calculated from the precision $\frac{\tp}{\tp + \fp}$ and the recall $\frac{\tp}{\tp + \fn}$ for each individual class $i$.
+Where $\tp$ is the true positives, $\fp$ is the false positives and $\fn$ is the false negatives.
+
+\begin{equation}\label{eq:mf1}
+    \text{F1}_{micro} = \frac{1}{N}\sum_{i=1}^N\frac{2 \cdot \tp_i}{2 \cdot \tp_i + \fp_i + \fn_i}
+\end{equation}

### 2024-05-28 paper/sections/04-method.tex
general small updates

@@
 \subsubsection{Principal Component Analysis}
-Principal Component Analysis (PCA) is a method for reducing dimensionality
+Principal Component Analysis (PCA) is a method for reducing dimensionality~\cite{pcapaper}.
+It functions by projecting the high dimensional data onto the direction of maximum variance, thus retaining key features and reducing noise.
 

### 2024-05-28 paper/sections/03-data.tex
Orgainised data section

@@
 
-Dataset Description
+The training data comprises species observations and environmental data. Below, we explain the data in detail.
 
-The training data comprises species observations and environmental data. Below, we explain the data in detail.
+\subsection{Observations data}
 
@@
 The presence-absence data (PA) is provided to compensate for the problem of false-absences of PO data and calibrate models to avoid associated biases.
+
 Presence-Only (PO) occurrences: combines around five million observations from numerous datasets gathered from the Global Biodiversity Information Facility (GBIF, www.gbif.org).
@@
 
-Environmental data
+\subsection{Environmental data}
 
@@
 
+\subsubsection{Satellite image patches}
+
 Satellite image patches: 3-band (RGB) and 1-band (NIR) 128x128 JPEG images, a color JPEG file for RGB data and a grayscale one for Near-Infrared images at 10m resolution.
 The source for these images is Sentinel2 remote sensing data pre-processed by the Ecodatacube platform.
+
+\subsubsection{Satellite time series}
 
@@
 The source for this is the Landsat remote sensing data pre-processed by the Ecodatacube platform
+
+\subsubsection{Environmental rasters}
 
@@
 
-Bioclimatic rasters: 19 low-resolution rasters covering Europe; commonly used in species distribution modeling.
+\begin{enumerate}
+    \item Bioclimatic rasters: 19 low-resolution rasters covering Europe; commonly used in species distribution modeling.
 Provided in longitude/latitude coordinates (WGS84).
@@
 The source for these rasters is the CHELSA climate dataset.
-
-Soil rasters: Nine pedologic low-resolution rasters covering Europe.
+    \item Soil rasters: Nine pedologic low-resolution rasters covering Europe.
 Provided variables describe the soil properties from 5 to 15cm depth and are determinant of plant species distributions.
@@
 The source for these rasters is Soilgrids.
-
-Elevation: High-resolution raster covering Europe.
+    \item Elevation: High-resolution raster covering Europe.
 Provided as a GeoTIFF file and CSV file with extracted values, with a resolution of 1 arc second (~30 meters).
 The source for this raster is the ASTER Global Digital Elevation Model V3.
-
-Land Cover: A medium-resolution multi-band land cover raster covering Europe.
+    \item Land Cover: A medium-resolution multi-band land cover raster covering Europe.
 Each band describes either the land cover class prediction or its confidence under various classifications.
@@
 The source for this raster is MODIS Terra+Aqua 500m.
-
-Human footprint: Several low-resolution rasters describing human footprint, encapsulating seven pressures on the environment (e.g., nighlight level, population density) induced by human presence and activity, are provided for two time periods, the early 90's (~1993) and late 2000' (~2009).
+    \item Human footprint: Several low-resolution rasters describing human footprint, encapsulating seven pressures on the environment (e.g., nighlight level, population density) induced by human presence and activity, are provided for two time periods, the early 90's (~1993) and late 2000' (~2009).
 We provide two summary rasters combining all human pressures and two detailed rasters per pressure, which avoid an arbitrary degradation of the original data.
@@
 The source for these rasters is \cite{human-footprint-dataset-2016}.
+\end{enumerate}

### 2024-05-28 paper/sections/02-background.tex

\section{Background}

The GeoLifeCLEF challenge has been running for a number of years. Each year, participants are tasked with predicting species distribution, but the challenge has evolved over time, with new datasets, evaluation metrics, and research questions introduced each year. Here, we provide an overview of the some of the recent submissions to GeoLifeCLEF challenge and summarize the key contributions.

In 2021, the GeoLifeCLEF challenge focused on fine-grained visual categorization using remote sensing data. The winning submission by \cite{Seneviratne:CLEF-2021} leveraged contrastive learning to improve species distribution modeling (SDM) from remote sensing imagery. The authors explored the effectiveness of using only RGB imagery and the impact of adding altitude imagery to the model's performance. They introduced a new consistency-based model selection metric to enhance the model's generalization capabilities. The paper outlined potential areas for further research, including the impact of transformations and the utility of the consistency metric.

In 2022, the GeoLifeCLEF challenge shifted its focus to predicting species distribution across the U.S. and France using remote sensing data and other covariates. The second-place submission by \cite{KellenbergerDevis:CLEF2022} proposed a classification approach with a spatial block-label swap regularization during training and an ensemble of deep learning models. Their method achieved a top-30 accuracy of 31.22\% on the private test set, securing second place in the competition. The authors reflected on the results and suggested potential improvements and the importance of species distribution modeling for ecological research.

In 2023, the GeoLifeCLEF challenge introduced a new dataset with single positive labels for each location, making multi-label prediction challenging. The winning submission by \cite{UngKojimaWada:CLEF2023} proposed a three-step training strategy to leverage the single positive labels effectively. The authors introduced several CNN-based models and demonstrated their effectiveness compared to a simple baseline. The paper discussed the challenges of the new dataset and the proposed models' performance, providing detailed results and comparisons.

### 2024-05-28 paper/sections/01-intro.tex
Add kaggle information to the introduction

@@
 
-\begin{figure}
-    \begin{center}
-        \input{figs/model-diagram.tex}
-        \caption{This diagram is just here temporarily and will be in a more relevant section when complete}
-    \end{center}
-\end{figure}
+% mention that the research is from a kaggle competition
+The research was conducted as part of the GeoLifeCLEF 2024 competition on Kaggle \cite{geolifeclef-2024}, which is a part of the LifeCLEF initiative.
+The competition aims to develop models for predicting plant species in a specific location and time using various environmental factors as predictors.

### 2024-05-29 paper/sections/04-method.tex
made preprocessing diagram
Add vision transformer section

@@
 For the purposes of this research we focused on ResNet18, the 18 layer deep variation with some modifications to accomodate the shape of the input data and the required output.
-\subsubsection{[Transformer]}
+\subsubsection{Transformer}
+Vision transformers (ViTs) have emerged as a novel approach to image classification, often outperforming traditional convolutional neural networks (CNNs) by leveraging self-attention mechanisms originally developed for machine translation \cite{vaswani2017attention}.
+Transformer now underpin some of the most powerful Large Language models \cite{brown2020language}.
+By tokenizing images into fixed-size patches and encoding them into embeddings for transformer layers to process, ViTs excel at capturing long-range dependencies within an image, enabling a more holistic interpretation of visual context compared to CNN's local focus \cite{dosovitskiy2021image}.
+This methodological shift allows the models to effectively decipher complex scenes and interactions that are typically challenging for traditional architectures.
+
+Often ViT-based models are often developed for large high resolution imagery.
+This was not the case for this application where the satellite images were a modest 128x128 pixels.
+We believed that the ability of transformer based models to capture to intricate regional relationships.
+Would be beneficial to the task of species classification.
 
@@
 
+\subsubsection{Preprocessing}
+
+\begin{figure}\label{fig:preproc}
+    \begin{center}
+        \input{figs/pca-diagram.tex}
+        \caption{Preprocessing the data for use in the count prediction and XGBoost pseudo-probability models}
+    \end{center}
+\end{figure}
+
+\subsubsection{Main Process}
 \newcommand{\nspecies}{N_{\text{species}}}
@@
 
-\subsubsection{Species scoring}
-\begin{figure}
-    \begin{center}
-        \input{figs/model-diagram.tex}
-        \caption{The structure of the scoring ensemble from data through to individual species score outputs.}
-    \end{center}\label{fig:score-structure}
-\end{figure}
-\subsubsection{Species counts}
-\begin{figure}
-    \begin{center}
-        \input{figs/count-diagram.tex}
-        \caption{The counts ensemble through to final outputs.}
-    \end{center}\label{fig:count-structure}
-\end{figure}
-\subsection{Metrics}
+% \subsubsection{Species scoring}
+% \begin{figure}
+%     \begin{center}
+%         \input{figs/model-diagram.tex}
+%         \caption{The structure of the scoring ensemble from data through to individual species score outputs.}
+%     \end{center}\label{fig:score-structure}
+% \end{figure}
+% \subsubsection{Species counts}
+% \begin{figure}
+%     \begin{center}
+%         \input{figs/count-diagram.tex}
+%         \caption{The counts ensemble through to final outputs.}
+%     \end{center}\label{fig:count-structure}
+% \end{figure}
+% \subsection{Metrics}
 

### 2024-06-04 models.py
Add further work to ViT

@@ -1,0 +1,108 @@
+import torchvision.models as tvm
+import torch.nn as nn
+import torch.nn.functional as F
+from einops import rearrange
+
+
+class MLP(nn.Module):
+    def __init__(self, dim, out_dim):
+        super().__init__()
+        self.fc1 = nn.Linear(dim, out_dim * 5)
+        self.fc2 = nn.Linear(out_dim * 5, out_dim)
+        self.norm = nn.LayerNorm(out_dim * 5)
+
+    def forward(self, x):
+        x = F.tanh(self.fc1(x))
+        x = self.norm(x)
+        x = self.fc2(x)
+        return x
+
+
+class Multihead(nn.Module):
+    def __init__(self, heads, head_dim, dim):
+        super().__init__()
+        self.head_dim = head_dim
+        self.heads = heads
+        self.inner_dim = self.heads * self.head_dim
+        self.scale = self.head_dim**-0.5
+        self.to_qkv = nn.Linear(dim, self.inner_dim * 3)
+        self.to_output = nn.Linear(self.inner_dim, dim)
+        self.norm = nn.LayerNorm(dim)
+        self.softmax = nn.Softmax(dim=-1)
+
+    def forward(self, x):
+        x = self.norm(x)
+        qkv = self.to_qkv(x).chunk(3, dim=-1)
+        Q, K, V = map(
+            lambda t: rearrange(t, "b l (h dim) -> b h l dim", dim=self.head_dim), qkv
+        )
+        K_T = K.transpose(-1, -2)
+        att_score = Q @ K_T * self.scale
+        att = self.softmax(att_score)
+        out = att @ V  # (B,H,L,dim)
+        out = rearrange(out, "b h l dim -> b l (h dim)")
+        output = self.to_output(out)
+        return output
+
+
+class FeedForward(nn.Module):
+    def __init__(self, dim, mlp_dim):
+        super().__init__()
+        self.fc1 = nn.Linear(dim, mlp_dim)
+        self.fc2 = nn.Linear(mlp_dim, dim)
+        self.norm = nn.LayerNorm(dim)
+
+    def forward(self, x):
+        x = self.norm(x)
+        x = F.gelu(self.fc1(x))
+        x = self.fc2(x)
+        return x
+
+
+class Transformer(nn.Module):
+    def __init__(self, dim, heads, head_dim, out_dim):
+        super().__init__()
[…]

### 2024-06-04 models2.py

@@ -1,0 +1,108 @@
+import torchvision.models as tvm
+import torch.nn as nn
+import torch.nn.functional as F
+from einops import rearrange
+
+
+class MLP(nn.Module):
+    def __init__(self, dim, out_dim):
+        super().__init__()
+        self.fc1 = nn.Linear(dim, out_dim * 5)
+        self.fc2 = nn.Linear(out_dim * 5, out_dim)
+        self.norm = nn.LayerNorm(out_dim * 5)
+
+    def forward(self, x):
+        x = F.tanh(self.fc1(x))
+        x = self.norm(x)
+        x = self.fc2(x)
+        return x
+
+
+class Multihead(nn.Module):
+    def __init__(self, heads, head_dim, dim):
+        super().__init__()
+        self.head_dim = head_dim
+        self.heads = heads
+        self.inner_dim = self.heads * self.head_dim
+        self.scale = self.head_dim**-0.5
+        self.to_qkv = nn.Linear(dim, self.inner_dim * 3)
+        self.to_output = nn.Linear(self.inner_dim, dim)
+        self.norm = nn.LayerNorm(dim)
+        self.softmax = nn.Softmax(dim=-1)
+
+    def forward(self, x):
+        x = self.norm(x)
+        qkv = self.to_qkv(x).chunk(3, dim=-1)
+        Q, K, V = map(
+            lambda t: rearrange(t, "b l (h dim) -> b h l dim", dim=self.head_dim), qkv
+        )
+        K_T = K.transpose(-1, -2)
+        att_score = Q @ K_T * self.scale
+        att = self.softmax(att_score)
+        out = att @ V  # (B,H,L,dim)
+        out = rearrange(out, "b h l dim -> b l (h dim)")
+        output = self.to_output(out)
+        return output
+
+
+class FeedForward(nn.Module):
+    def __init__(self, dim, mlp_dim):
+        super().__init__()
+        self.fc1 = nn.Linear(dim, mlp_dim)
+        self.fc2 = nn.Linear(mlp_dim, dim)
+        self.norm = nn.LayerNorm(dim)
+
+    def forward(self, x):
+        x = self.norm(x)
+        x = F.gelu(self.fc1(x))
+        x = self.fc2(x)
+        return x
+
+
+class Transformer(nn.Module):
+    def __init__(self, dim, heads, head_dim, out_dim):
+        super().__init__()
[…]

### 2024-06-04 paper/sections/04-method.tex
finished new pipeline and data processor. Run preprocess_save.py

--- 2024-05-28 paper/sections/04-method.tex
@@
-\clearpage  % TODO comment this out when finished
-
 \section{Method}
@@
 
+\subsubsection{Preprocessing}
+
+The raw data included some missing, or infinite values that needed to be processed prior to model fitting.
+When a column was both deemed important to model fitting and contained such values, these values were replaced by the mean so as to avoid excessive influence from outliers without overly effecting the shape of the data~\cite{missingmedian}.
+
+The Metadata and Bioclimatic Rasters were preprocessed for use in the count and XGBoost models in order to both reduce dimensionality and remove potential noise.
+
+In addition, the number of viable output species was reduced from 11255 (the total species present accross all data) down to 1141 (the total species with more than 100 occurrences in the presence/absence data).
+This reduction served to remove several edge cases and focus the models more on likelier species.
+
+\begin{figure}\label{fig:preproc}
+    \begin{center}
+        \input{figs/pca-diagram.tex}
+        \caption{Preprocessing the data for use in the count prediction and XGBoost pseudo-probability models}
+    \end{center}
+\end{figure}
+
+\subsubsection{Main Process}
 \newcommand{\nspecies}{N_{\text{species}}}

### 2024-06-04 pipeline2/__init__.py

@@ -1,0 +1,2 @@
+from .preprocessing.preprocessor import GeoLifePreprocessor
+from .preprocessing.pcamanager import GeoLifePCA

### 2024-06-04 pipeline2/preprocessing/preprocessor.py

@@ -1,0 +1,159 @@
+# from pipeline.preprocess import GeoLifePreprocessor
+import os 
+import numpy as np 
+import pandas as pd
+
+class GeoLifePreprocessor():
+    """ Collects the raw data from the Metadata and Environmental Rasters into a
+    single set of master data frames
+    """
+    __drop_list = [         # Occurences after group by survey:
+        "PANNONIAN",            # 248 occurences 
+        "STEPPIC",              # 48 occurences 
+        "BLACK SEA",            # 8 occurences
+        "Hungary",              # 29 occurences
+        "Ireland",              # 12 occurences
+        "Monaco",               # 7 occurences
+        "The former Yugoslav Republic of Macedonia", # 6 occurences
+        "Portugal",             # 6 occurences
+        "Norway",               # 5 occurences
+        "Andorra",              # 4 occurences
+        "Luxembourg",           # 2 occurences
+    ]
+    __reassign_list = [
+        "Bosnia and Herzegovina",
+        "Bulgaria",
+        "Slovenia",
+        "Greece",
+        "Serbia",
+        "Slovakia",
+        "Croatia",
+    ]
+    _rasters = None 
+    _data = None
+
+    def __init__(self, base_path:str = "data/_raw") -> None:
+        self._base_path = base_path 
+        self._raster_path = os.path.join(
+            self._base_path, 
+            "EnvironmentalRasters",
+            "EnvironmentalRasters"
+        )
+
+
+    def process_data(self) -> None: 
+        """ Loads in and processes the data for both test and train datasets
+        """
+        train = None 
+        test = None
+        for purpose in ["train", "test"]:
+            self._combine_rasters(is_train=(purpose == "train"))
+            self._get_metadata(is_train=(purpose == "train"))
+            self._encode_data(is_train=(purpose == "train"))
+            if purpose == "train":
+                train = self._data.join(self._rasters)
+                print("Train done")
+            else:
+                test = self._data.join(self._rasters)
+                print("Test done")
+        return train, test
+
+    def get_data(self):
+        if self._data is not None:
+            return self._data 
+        print("Data is not yet processed. Run process_data() and try again.")
+
+    def _combine_rasters(self, is_train:bool) -> None:
+        """
+        """
+        self._rasters = None 
+        for raster in [
+            "Elevation",
+            "Human Footprint",
+            "LandCover",
+            "SoilGrids",
+        ]:
+            loaded = pd.read_csv(
+                os.path.join(
+                    self._raster_path,
[…]

### 2024-06-04 paper/sections/04-method.tex
updates to paper files
resolved merge conflict in bibliography

--- 2024-05-29 paper/sections/04-method.tex
@@
 \subsubsection{Preprocessing}
+
+The raw data included some missing, or infinite values that needed to be processed prior to model fitting.
+When a column was both deemed important to model fitting and contained such values, these values were replaced by the mean so as to avoid excessive influence from outliers without overly effecting the shape of the data~\cite{missingmedian}.
+
+\textbf{TODO: Further explanation when models finalized}
+
+The Metadata and Bioclimatic Rasters were preprocessed for use in the count and XGBoost models in order to both reduce dimensionality and remove potential noise.
+
+In addition, the number of viable output species was reduced from 11255 (the total species present accross all data) down to 1141 (the total species with more than 100 occurrences in the presence/absence data).
+This reduction served to remove several edge cases and focus the models more on likelier species.
 

### 2024-06-06 pipeline2/__init__.py
XGBoost model and preprocessing complete

@@ -1,2 +1,6 @@
 from .preprocessing.preprocessor import GeoLifePreprocessor
 from .preprocessing.pcamanager import GeoLifePCA
+
+from .model_management.xgb_manager import GeoLifeXGB
+
+from .postprocessing.postprocessor import GeoLifePostprocessor

### 2024-06-06 pipeline2/model_management/xgb_manager.py

@@ -1,0 +1,77 @@
+import numpy as np
+import pickle as pk
+
+from xgboost import XGBRegressor 
+
+
+class GeoLifeXGB():
+
+    __main_params = {
+        'device'        : 'cuda', 
+        'reg_lambda'    : 12, 
+        'learning_rate' : 0.1, 
+        'min_split_loss': 0, 
+        'reg_alpha'     : 0
+    }
+
+    __count_params = {
+        'device'        : 'cuda',
+        'learning_rate' : 0.1,
+        'min_split_loss': 0,
+        'reg_lambda'    : 10,
+        'reg_alpha'     : 0, 
+    }
+
+    count_model = None
+    main_model = None
+
+    def __init__(self, seed) -> None:
+        self.seed = seed 
+        self.__load_data()
+
+    def generate_count_model(self):
+        self.count_model = XGBRegressor(**self.__count_params, seed=self.seed)
+        self.count_model.fit(
+            self.train.drop(columns=self.yy_cols),
+            self.train[self.yy_cols].apply(lambda x: x.sum(), axis=1)
+            
+        )
+        return self.count_model
+    
+    def predict_test_counts(self, addition:float = 3.5):
+        if self.count_model is None:
+            print("Train or load a model before predicting.")
+            return 
+        
+        counts = self.count_model.predict(self.test) + addition 
+        return counts.round().astype(int)
+
+    def generate_main_model(self):
+        self.main_model = XGBRegressor(**self.__main_params, seed=self.seed)
+        combined_train = self.train.drop(columns=self.yy_cols)
+        combined_train[
+            ["landsat" + str(idx) for idx in range(360)]
+        ] = self.landsat
+        self.main_model.fit(combined_train, self.train[self.yy_cols])
+        return self.main_model 
+    
+    def predict_test_scores(self):
+        if self.main_model is None:
+            print("Train or load a model before predicting.")
+        combined_test = self.test 
+        combined_test[
+            ["landsat" + str(idx) for idx in range(360)]
+        ] = self.landsat_test 
+        return self.main_model.predict(combined_test)
+ 
+    def __load_data(self):
+        self.train = pk.load(open("data/processed/train.pkl", "rb"))
+        self.test = pk.load(open("data/processed/test.pkl", "rb"))
+        self.yy_cols = [col for col in self.train if "speciesId_" in col]
+        self.landsat = pk.load(open("data/processed/train_ls.pkl", "rb"))
[…]

### 2024-06-06 pipeline2/postprocessing/applycounts.py

@@ -1,0 +1,14 @@
+import numpy as np
+
+def apply_counts(predictions, counts):
+    final_output = np.zeros(predictions.shape, dtype=bool)
+    
+    for idx in range(predictions.shape[0]):
+        for col in np.argpartition(
+            predictions[idx, :], 
+            -counts[idx]
+        )[-counts[idx]:]:
+            final_output[idx, col] = 1 
+    
+    return final_output
+        

### 2024-06-06 pipeline2/preprocessing/landsat.py

@@ -1,0 +1,51 @@
+import numpy as np
+import pandas as pd
+import pickle as pk
+
+TRAIN_IDS = pk.load(open("data/processed/train.pkl", "rb")).index
+TRAIN_IDS = TRAIN_IDS.to_list()
+COLORS = ["red", "green", "blue", "nir", "swir1", "swir2"]
+
+def get_path(is_train, color):
+    is_train = "train" if is_train else "test"
+    path = f"data/_raw/PA-{is_train}-landsat_time_series/GLC24-PA-{is_train}"
+    path += f"-landsat_time_series-{color}.csv"
+    return path 
+
+def pickle_landsat():
+    for truth in [True, False,]:
+        image_data = None 
+        for c_idx, color in enumerate(COLORS):
+            path = get_path(truth, color)
+            data = pd.read_csv(path)
+            if image_data is None:
+                axis_0 = len(TRAIN_IDS) if truth else data.shape[0]
+                image_data = np.zeros((axis_0, 60, 6), np.uint8)
+            cols = data.columns.to_list()
+            cols.reverse()
+            data = data[cols]
+            data["from"] = data.apply(
+                lambda x: str(int(x.first_valid_index()[:4]) - 15) + "_1",
+                axis=1
+            )
+            data.reset_index()
+            idx = 0
+            for _, row in data.iterrows():
+                if truth and row["surveyId"] not in TRAIN_IDS: 
+                    continue
+                
+                for modifier in range(0, 15):
+                    for quarter in range(1, 5):
+                        image_data[
+                            idx, 4 * modifier + quarter - 1, c_idx
+                        ] = np.uint8(
+                            row[f"{int(row['from'][:4]) + modifier}_{quarter}"]
+                        )
+                idx += 1
+        save_path = f"data/processed/full_{'train' if truth else 'test'}"
+        save_path += "_ls.pkl"
+        pk.dump(image_data, open(save_path, "wb"))
+
+
+if __name__ == "__main__":
+    pickle_landsat()

### 2024-06-06 pipeline2/preprocessing/preprocessor.py

@@ -40,7 +40,6 @@ class GeoLifePreprocessor():
             "EnvironmentalRasters"
         )
 
-
     def process_data(self) -> None: 
         """ Loads in and processes the data for both test and train datasets
         """
@@ -50,6 +49,8 @@ class GeoLifePreprocessor():
             self._combine_rasters(is_train=(purpose == "train"))
             self._get_metadata(is_train=(purpose == "train"))
             self._encode_data(is_train=(purpose == "train"))
+            self._data.replace([np.inf, -np.inf], np.nan, inplace=True)
+            self._data = self._data.fillna(self._data.median())
             if purpose == "train":
                 train = self._data.join(self._rasters)
                 print("Train done")
@@ -57,11 +58,6 @@ class GeoLifePreprocessor():
                 test = self._data.join(self._rasters)
                 print("Test done")
         return train, test
-
-    def get_data(self):
-        if self._data is not None:
-            return self._data 
-        print("Data is not yet processed. Run process_data() and try again.")
 
     def _combine_rasters(self, is_train:bool) -> None:
         """
@@ -101,6 +97,7 @@ class GeoLifePreprocessor():
                 os.path.join(self._base_path, "GLC24_PA_metadata_test.csv"),
                 index_col="surveyId"
             )
+
             return 
         
         self._data = pd.read_csv(

### 2024-06-06 pipeline2/preprocessing/landsat.py
finalised xgb

@@ -42,7 +42,7 @@ def pickle_landsat():
                             row[f"{int(row['from'][:4]) + modifier}_{quarter}"]
                         )
                 idx += 1
-        save_path = f"data/processed/full_{'train' if truth else 'test'}"
+        save_path = f"data/processed/{'train' if truth else 'test'}"
         save_path += "_ls.pkl"
         pk.dump(image_data, open(save_path, "wb"))
 

### 2024-06-06 preprocess_data.py
added preprocess file

@@ -1,8 +1,12 @@
 import pickle as pk
 from pipeline2 import GeoLifePreprocessor, GeoLifePCA
+from pipeline2.preprocessing import pickle_landsat
 
 if __name__ == "__main__":
     train, test = GeoLifePreprocessor().process_data()
     train, test = GeoLifePCA(train, test).process_data()
     pk.dump(train, open("data/processed/train.pkl", "wb"))
     pk.dump(test, open("data/processed/test.pkl", "wb"))
+    train_ls, test_ls = pickle_landsat()
+    pk.dump(train_ls, open("data/processed/train_ls", "wb"))
+    pk.dump(test_ls, open("data/processed/test_ls", "wb"))

### 2024-06-06 pipeline2/preprocessing/__init__.py
added missing init file to pipeline

@@ -1,0 +1,1 @@
+from .landsat import pickle_landsat

### 2024-06-06 ensemblexgb.py
removed ensemblexgb

@@ -1,127 +1,0 @@
-""" python ensemblexgb.py [--train --load <model-name> <count-name> --run]
-"""
-import sys
-
-from xgboost import XGBRegressor
-
-from pipeline.preprocess import PCAApplier
-from pipeline.modelprocessing import save_model, load_model
-from pipeline.postprocessing import Postprocessor
-from helper.top_n_regression import get_top_n
-
-SEED = 31
-
-class XGBPipeline():
-    """ Creates and trains, or loads in XGBoost models and runs the training 
-    data through them
-    """
-
-    model = None
-    count_model = None
-    column_labels = None
-    model_type = "xgbregressor"
-
-    def train_models(self, device="cuda"):
-        """ Train the XGBoost models. Loads in the training data, creates the
-        main and count models and then fits them and saves them.
-
-        args:
-            device      :   "cuda" to use gpu, else "cpu" to use the cpu
-        """
-        train = PCAApplier('data/processed/train_full.pkl', method='full').data
-        xx_cols = [col for col in train.columns if "species" not in col]
-        yy_cols = [col for col in train.columns if "species" in col]
-        self.model = XGBRegressor(device=device)
-        self.count_model = XGBRegressor(device=device)
-        print(":: Fitting main model")
-        self.model.fit(train[xx_cols], train[yy_cols])
-        save_model(self.model, "xgbregressor", "-full-main", yy_cols)
-        print(":: Fitting species count model")
-        self.count_model.fit(
-            train[xx_cols],
-            train[yy_cols].apply(lambda x: x.sum(), axis=1)
-        )
-        save_model(self.count_model, self.model_type, "-full-counts")
-        self.column_labels = yy_cols
-
-    def load_models(self, model_name, count_name):
-        """ Loads in the pretrained models and stores them in member variables
-        note: the filenames work with or without the .pkl extension, the 
-        directory path is not needed
-        
-        args:
-            model_name  :   The file name of the main model
-            count_name  :   The file name of the count model
-        """
-        print(f":: Loading {model_name} as main model")
-        self.model = load_model(self.model_type, model_name)
-        print(f":: Loading {model_name} as species count model")
-        self.count_model = load_model(self.model_type, count_name)
-        self.column_labels = load_model(self.model_type, f"{model_name}_cols")
-
-
-    def run(self, additional_count: int = None):
-        """ Runs the models on the test data. Processes the data into uploadable
-        format and saves it in the submissions directory
-
-        args:
[…]

### 2024-06-06 helper/__init__.py

@@ -1,7 +1,7 @@
 """ import helper
 """
-from .verifytest import verify_test
-from .verifysubmission import verify_submission
-from .images_to_np import load_images
-from .tifnormalizer import tif_helper
+# from .verifytest import verify_test
+# from .verifysubmission import verify_submission
+# from .images_to_np import load_images
+# from .tifnormalizer import tif_helper
 from .top_n_regression import get_top_n

### 2024-06-06 models.py
tidied up files

@@ -1,108 +1,0 @@
-import torchvision.models as tvm
-import torch.nn as nn
-import torch.nn.functional as F
-from einops import rearrange
-
-
-class MLP(nn.Module):
-    def __init__(self, dim, out_dim):
-        super().__init__()
-        self.fc1 = nn.Linear(dim, out_dim * 5)
-        self.fc2 = nn.Linear(out_dim * 5, out_dim)
-        self.norm = nn.LayerNorm(out_dim * 5)
-
-    def forward(self, x):
-        x = F.tanh(self.fc1(x))
-        x = self.norm(x)
-        x = self.fc2(x)
-        return x
-
-
-class Multihead(nn.Module):
-    def __init__(self, heads, head_dim, dim):
-        super().__init__()
-        self.head_dim = head_dim
-        self.heads = heads
-        self.inner_dim = self.heads * self.head_dim
-        self.scale = self.head_dim**-0.5
-        self.to_qkv = nn.Linear(dim, self.inner_dim * 3)
-        self.to_output = nn.Linear(self.inner_dim, dim)
-        self.norm = nn.LayerNorm(dim)
-        self.softmax = nn.Softmax(dim=-1)
-
-    def forward(self, x):
-        x = self.norm(x)
-        qkv = self.to_qkv(x).chunk(3, dim=-1)
-        Q, K, V = map(
-            lambda t: rearrange(t, "b l (h dim) -> b h l dim", dim=self.head_dim), qkv
-        )
-        K_T = K.transpose(-1, -2)
-        att_score = Q @ K_T * self.scale
-        att = self.softmax(att_score)
-        out = att @ V  # (B,H,L,dim)
-        out = rearrange(out, "b h l dim -> b l (h dim)")
-        output = self.to_output(out)
-        return output
-
-
-class FeedForward(nn.Module):
-    def __init__(self, dim, mlp_dim):
-        super().__init__()
-        self.fc1 = nn.Linear(dim, mlp_dim)
-        self.fc2 = nn.Linear(mlp_dim, dim)
-        self.norm = nn.LayerNorm(dim)
-
-    def forward(self, x):
-        x = self.norm(x)
-        x = F.gelu(self.fc1(x))
-        x = self.fc2(x)
-        return x
-
-
-class Transformer(nn.Module):
-    def __init__(self, dim, heads, head_dim, out_dim):
-        super().__init__()
[…]

### 2024-06-06 setup.py

@@ -1,27 +1,0 @@
-#!/usr/bin/env python
-"""Downloads and preprocesses the kaggle data
-"""
-import sys
-from pipeline.getdata import DirectoryManagement
-
-# from pipeline.preprocess import pickle_landsat
-from pipeline.preprocess import PCACreator
-from pipeline.preprocess import DataCollector
-
-if __name__ == "__main__":
-    args = sys.argv
-    if "-h" in args or len(args) == 1:
-        print(
-            ":: Flags:\n\t-d  :  Download the data from kaggle (requires "
-            + "kaggle CLI)\n\t-p  :  Preprocess and pickle the data as single "
-            + "train and test files\n\t-r  :  Reduce the dimensions of the "
-            + "data using PCA, saving the PCA models for later use"
-        )
-    if "-d" in args:
-        DirectoryManagement()
-    if "-p" in args:
-        DataCollector().process_and_save_data()
-    if "-r" in args:
-        PCACreator(method="full")
-    if "-l" in args:
-        pickle_landsat()

### 2024-06-06 archive/data-top500.py
removed archive

@@ -1,148 +1,0 @@
-# Data Processor
-#
-# Desired output
-# - Full train dataset - for training model for Kaggle submission
-# - Full test dataset - for Kaggle submission
-# - Internal train dataset - 80% of full train dataset for internal banchmarking
-# - Internal test dataset - 20% of full train dataset for internal benchmarking
-# - All tabular data in one file
-
-import pandas as pd
-import numpy as np
-from tqdm import tqdm
-import shutil
-from pathlib import Path
-
-print("Processing data\n")
-
-# Load data
-print("Load data")
-train_df = pd.read_csv("data/GLC24_PA_metadata_train.csv")
-
-# Drop: 'Andorra', 'Hungary', 'Ireland', 'Latvia', 'Luxembourg', 'Monaco', 'Norway',
-# 'Portugal', 'Romania', 'Serbia', 'The former Yugoslav Republic of Macedonia' they
-# are not represented in the test set
-print("Drop countries not in test data")
-drop_countries = {
-    "Andorra",
-    "Hungary",
-    "Ireland",
-    "Latvia",
-    "Luxembourg",
-    "Monaco",
-    "Norway",
-    "Portugal",
-    "Romania",
-    "Serbia",
-    "The former Yugoslav Republic of Macedonia",
-}
-train_df = train_df[~train_df.country.isin(drop_countries)]
-
-# drop species outside top 500
-print("Drop species outside top 500")
-species = train_df.speciesId.value_counts().head(500).index
-train_df = train_df[train_df.speciesId.isin(species)]
-
-# -inf, inf replaced by NaN
-print("Replace -inf, inf with NaN")
-train_df = train_df.replace([np.inf, -np.inf], np.nan)
-
-# For the train data NaN in geoUncertaintyInM, and areaInM2 replaced with the country median values
-print("Process NaN values")
-for column in ["areaInM2", "geoUncertaintyInM"]:
-    print(f" - Processing train data: {column}")
-    for country in tqdm(train_df.country.unique()):
-        train_df.loc[train_df.country == country, column] = train_df.query(
-            f"country == '{country}'"
-        )[column].fillna(train_df.query(f"country == '{country}'")[column].median())
-
-
-# Resulting dataframes
-print("Resulting dataframes")
-print(f" - Train data: {train_df.shape}")
-
-# set speciesId as int
-print("Set speciesId as int")
-train_df["speciesId"] = train_df["speciesId"].astype(int)
-
-# Resulting dataframes
[…]

### 2024-06-06 archive/imageloader.py

@@ -1,47 +1,0 @@
-import numpy as np
-import cv2
-
-class ImageLoader():
-
-    def _get_image_from_id(self, id: int, rgb: bool, train: bool, 
-                                                data_path: str = "data/"):
-        """
-        """
-        file_location = data_path + self.__get_dir(train, rgb) \
-                      + self.__parse_img_id(id)
-        try:
-            image_array = cv2.imread(file_location)
-        except (FileNotFoundError):
-            message = "The generated file path does not correspond with an " \
-                    + "extant image file\n::\t ID:\t" + id + "\n::\t Path\t" \
-                    + self.__parse_img_id(id)
-            Exception(message)
-        return image_array
-        
-    def __get_dir(self, is_train: bool = True, is_rgb: bool = True):
-        t_or_t = "Train" if is_train else "Test"
-        r_or_n = "RGB" if is_rgb else "NIR"
-        path = "PA_" + t_or_t + "_SatellitePatches_" + r_or_n + "/pa_" \
-             + t_or_t.lower() + "_patches_" + r_or_n.lower() + '/'
-        return path       
-
-    def __parse_img_id(self, id: int):
-        id = str(id)
-
-        match len(id):
-            case 1:
-                ab = 1
-                cd = id 
-            case 2:
-                ab = 1
-                cd = id 
-            case 3:
-                ab = id[-3]
-                cd = id[-2:]
-            case _:
-                ab = id[-4:-2]
-                cd = id[-2:]
-
-        paresed_id = '/'.join([str(cd), str(ab), str(id)])
-        return paresed_id + ".jpeg"
-

### 2024-06-06 archive/top500_image.py

@@ -1,26 +1,0 @@
-
-import numpy as np
-import pandas as pd
-from PIL import Image
-from tqdm import tqdm
-
-top500_df = pd.read_pickle("../processed_data/partially_processed/top500_df_elevation_onehot_grouped.pkl")
-
-images = []
-
-for survey_id in tqdm(top500_df.surveyId):
-    survey_id = str(survey_id)
-    cd = survey_id[-2:]
-    ab = survey_id[-4:-2]
-
-    ir_image_path = f'../data/PA_Train_SatellitePatches_NIR/pa_train_patches_nir/{cd}/{ab}/{survey_id}.jpeg'
-    pil_ir_img = Image.open(ir_image_path)
-
-    rgb_image_path = f'../data/PA_Train_SatellitePatches_RGB/pa_train_patches_rgb/{cd}/{ab}/{survey_id}.jpeg'
-    pil_rgb_img = Image.open(rgb_image_path)
-
-    _, b, g = pil_rgb_img.split()
-
-    images.append(np.asarray(Image.merge('RGB', (pil_ir_img, g, b))))
-
-np.savez_compressed("../processed_data/partially_processed/top500_images.npz", np.array(images))

### 2024-06-06 archive/top500_ohe.py

@@ -1,27 +1,0 @@
-
-import numpy as np
-import pandas as pd
-from sklearn.preprocessing import OneHotEncoder
-
-top500_df = pd.read_csv("../processed_data/partially_processed/top500.csv")
-
-top500_df = top500_df.drop(columns=['geoUncertaintyInM', 'country'])
-
-with pd.option_context('use_inf_as_na', True):
-    top500_df.areaInM2 = top500_df.areaInM2.fillna(top500_df.areaInM2.mean())
-
-top500_df.speciesId = top500_df.speciesId.astype('uint16')
-
-ohe = OneHotEncoder(dtype='bool')
-
-transformed = ohe.fit_transform(top500_df[['region', 'speciesId']])
-
-top500_df[np.concatenate((ohe.categories_[0], ohe.categories_[1]), axis=0)] = transformed.toarray()
-
-top500_df_grouped = top500_df.groupby('surveyId', as_index=False).max().copy()
-
-print(top500_df_grouped.shape)
-
-top500_df_grouped.to_pickle("../processed_data/partially_processed/top500_df_elevation_onehot_grouped.pkl")
-
-

### 2024-06-06 data-pipeline/1-top-n.py
removed data-pipeline

@@ -1,55 +1,0 @@
-import pandas as pd
-import argparse as ap
-import os
-from pathlib import Path
-
-
-def init_argparse() -> ap.ArgumentParser:
-    parser = ap.ArgumentParser(
-        usage="%(prog)s [OPTION]",
-        description="Filter the top N species from the training dataset.",
-    )
-    parser.add_argument(
-        "--number",
-        "-n",
-        type=int,
-        help="The number of species to retain.",
-        default=500,
-    )
-    parser.add_argument(
-        "--file",
-        "-f",
-        type=str,
-        help="The path to the training dataset.",
-        default="_raw/GLC24_PA_metadata_train.csv",
-    )
-    return parser
-
-
-args = init_argparse().parse_args()
-assert args.number > 0, f"Invalid number of species, got {args.number}."
-assert os.path.exists(args.file), f"File '{args.file}' not found."
-
-if args.file[-3:] == "csv":
-    train_df = pd.read_csv(args.file)
-elif args.file[-3:] == "pkl":
-    train_df = pd.read_pickle(args.file)
-else:
-    raise ValueError("Invalid file format, must be either CSV or PKL.")
-
-assert "speciesId" in train_df.columns, "Column 'speciesId' not found."
-
-top_n = sorted(
-    list(train_df.speciesId.value_counts().head(args.number).index.astype(int))
-)
-top_n_df = train_df[train_df.speciesId.isin(top_n)]
-
-top_n_df.loc[:, "region"] = top_n_df.region.astype("category")
-top_n_df.loc[:, "country"] = top_n_df.country.astype("category")
-top_n_df.loc[:, "speciesId"] = top_n_df.speciesId.astype(int)
-
-print(f"Original data: {train_df.shape}\nTop {args.number} species: {top_n_df.shape}")
-
-Path("output").mkdir(parents=True, exist_ok=True)
-top_n_df.to_pickle(f"output/1-top-{args.number}_train.pkl")
-print(f"1-top-{args.number}_train.pkl has shape {top_n_df.shape}")

### 2024-06-06 data-pipeline/10-finalise.py

@@ -1,98 +1,0 @@
-# Data Processor
-#
-# Desired output
-# - Full train dataset - for training model for Kaggle submission
-# - Full test dataset - for Kaggle submission
-# - Internal train dataset - 80% of full train dataset for internal banchmarking
-# - Internal test dataset - 20% of full train dataset for internal benchmarking
-# - All tabular data in one file
-
-import pandas as pd
-import shutil
-from tqdm import tqdm
-from pathlib import Path
-import argparse as ap
-import os
-
-
-def init_argparse() -> ap.ArgumentParser:
-    parser = ap.ArgumentParser(
-        usage="%(prog)s [OPTION]",
-        description="Convert the satellite images to PNG format.",
-    )
-    parser.add_argument(
-        "--train",
-        type=str,
-        help="The path to the training dataset.",
-        default="output/9-int-split-train.pkl",
-    )
-    parser.add_argument(
-        "--test",
-        type=str,
-        help="The path to the testing dataset.",
-        default="output/9-int-split-test.pkl",
-    )
-    parser.add_argument(
-        "--desc",
-        type=str,
-        help="A description of the data set to be used for the file path.",
-        default="final",
-    )
-    return parser
-
-
-args = init_argparse().parse_args()
-assert os.path.exists(args.train), f"Training data file '{args.train}' not found."
-assert os.path.exists(args.test), f"Testing data file '{args.test}' not found."
-
-if args.train[-3:].lower() == "csv":
-    train_df = pd.read_csv(args.train)
-elif args.train[-3:].lower() == "pkl":
-    train_df = pd.read_pickle(args.train)
-else:
-    raise ValueError("Invalid file format, must be either CSV or PKL.")
-
-if args.test[-3:].lower() == "csv":
-    test_df = pd.read_csv(args.test)
-elif args.test[-3:].lower() == "pkl":
-    test_df = pd.read_pickle(args.test)
-else:
-    raise ValueError("Invalid file format, must be either CSV or PKL.")
-
-test_X_df = test_df.loc[:, ~test_df.columns.str.startswith("speciesId_")]
-species_columns = ["surveyId"] + list(
-    test_df.columns[test_df.columns.str.startswith("speciesId_")]
-)
-test_y_df = test_df.loc[:, species_columns]
[…]

### 2024-06-06 data-pipeline/11-full-y-for-test-9.py

@@ -1,15 +1,0 @@
-import pandas as pd
-
-# Read in full species data and one hot encode it
-yy = pd.read_csv("_raw/GLC24_PA_metadata_train.csv", index_col="surveyId")
-yy = pd.get_dummies(yy["speciesId"].astype(int), prefix="species")
-
-# Load in the survey ID values for the test data
-test_surveys = list(pd.read_pickle("output/9-int-split-test.pkl")["surveyId"])
-
-# Strip out non-test surveys and group by remaining survey IDs
-yy = yy[yy.index.isin(test_surveys)]
-yy = yy.groupby(by=yy.index).max()
-
-# Pickle file
-yy.to_pickle("output/9-int-split-test-full-labels.pkl")

### 2024-06-06 data-pipeline/2-img-to-png.py

@@ -1,75 +1,0 @@
-from PIL import Image
-from tqdm import tqdm
-from pathlib import Path
-import pandas as pd
-import argparse as ap
-import os
-
-
-def init_argparse() -> ap.ArgumentParser:
-    parser = ap.ArgumentParser(
-        usage="%(prog)s [OPTION]",
-        description="Convert the satellite images to PNG format.",
-    )
-    parser.add_argument(
-        "--train",
-        type=str,
-        help="The path to the training dataset.",
-        default="output/1-top-500_train.pkl",
-    )
-    parser.add_argument(
-        "--test",
-        type=str,
-        help="The path to the testing dataset.",
-        default="_raw/GLC24_PA_metadata_test.csv",
-    )
-    parser.add_argument(
-        "--desc",
-        type=str,
-        help="A description of the data set to be used for the file path.",
-        default="all",
-    )
-    return parser
-
-
-args = init_argparse().parse_args()
-assert os.path.exists(args.train), f"Training data file '{args.train}' not found."
-assert os.path.exists(args.test), f"Testing data file '{args.test}' not found."
-
-if args.train[-3:].lower() == "csv":
-    train_survey_ids = pd.read_csv(args.train).surveyId.unique().astype(str)
-elif args.train[-3:].lower() == "pkl":
-    train_survey_ids = pd.read_pickle(args.train).surveyId.unique().astype(str)
-else:
-    raise ValueError("Invalid file format, must be either CSV or PKL.")
-
-if args.test[-3:].lower() == "csv":
-    test_survey_ids = pd.read_csv(args.test).surveyId.unique().astype(str)
-elif args.test[-3:].lower() == "pkl":
-    test_survey_ids = pd.read_pickle(args.test).surveyId.unique().astype(str)
-else:
-    raise ValueError("Invalid file format, must be either CSV or PKL.")
-
-train_path = Path(f"output/2-{args.desc}_images") / "train"
-test_path = Path(f"output/2-{args.desc}_images") / "test"
-Path(train_path).mkdir(parents=True, exist_ok=True)
-Path(test_path).mkdir(parents=True, exist_ok=True)
-
-for set_type, survey_ids, path in [
-    ("train", train_survey_ids, train_path),
-    ("test", test_survey_ids, test_path),
-]:
-    for survey_id in tqdm(survey_ids):
[…]

### 2024-06-06 data-pipeline/3-metadata-tidy.py

@@ -1,66 +1,0 @@
-from PIL import Image
-from tqdm import tqdm
-from pathlib import Path
-import pandas as pd
-import argparse as ap
-import os
-
-
-def init_argparse() -> ap.ArgumentParser:
-    parser = ap.ArgumentParser(
-        usage="%(prog)s [OPTION]",
-        description="Convert the satellite images to PNG format.",
-    )
-    parser.add_argument(
-        "--train",
-        type=str,
-        help="The path to the training dataset.",
-        default="output/1-top-500_train.pkl",
-    )
-    parser.add_argument(
-        "--desc",
-        type=str,
-        help="A description of the data set to be used for the file path.",
-        default="m-tidy",
-    )
-    return parser
-
-
-args = init_argparse().parse_args()
-assert os.path.exists(args.train), f"Training data file '{args.train}' not found."
-
-if args.train[-3:].lower() == "csv":
-    train_df = pd.read_csv(args.train)
-elif args.train[-3:].lower() == "pkl":
-    train_df = pd.read_pickle(args.train)
-else:
-    raise ValueError("Invalid file format, must be either CSV or PKL.")
-
-drop_countries = {
-    "Andorra",
-    "Hungary",
-    "Ireland",
-    "Latvia",
-    "Luxembourg",
-    "Monaco",
-    "Norway",
-    "Portugal",
-    "Romania",
-    "Serbia",
-    "The former Yugoslav Republic of Macedonia",
-}
-train_df = train_df[~train_df.country.isin(drop_countries)]
-
-drop_regions = {
-    "BLACK SEA",
-    "PANNONIAN",
-}
-train_df = train_df[~train_df.region.isin(drop_regions)]
-
-train_df.surveyId = train_df.surveyId.astype(int)
-train_df.speciesId = train_df.speciesId.astype(int)
-
-print(f"{train_df.shape=}")
-
-train_df.to_pickle(f"output/3-{args.desc}_train.pkl")
-print(f"3-{args.desc}_train.pkl has shape {train_df.shape}")

### 2024-06-06 data-pipeline/4-onehotencode.py

@@ -1,74 +1,0 @@
-from PIL import Image
-from tqdm import tqdm
-from pathlib import Path
-import pandas as pd
-import argparse as ap
-import os
-
-
-def init_argparse() -> ap.ArgumentParser:
-    parser = ap.ArgumentParser(
-        usage="%(prog)s [OPTION]",
-        description="Convert the satellite images to PNG format.",
-    )
-    parser.add_argument(
-        "--train",
-        type=str,
-        help="The path to the training dataset.",
-        default="output/3-m-tidy_train.pkl",
-    )
-    parser.add_argument(
-        "--test",
-        type=str,
-        help="The path to the testing dataset.",
-        default="_raw/GLC24_PA_metadata_test.csv",
-    )
-    parser.add_argument(
-        "--desc",
-        type=str,
-        help="A description of the data set to be used for the file path.",
-        default="ohe",
-    )
-    return parser
-
-
-args = init_argparse().parse_args()
-assert os.path.exists(args.train), f"Training data file '{args.train}' not found."
-assert os.path.exists(args.test), f"Testing data file '{args.test}' not found."
-
-if args.train[-3:].lower() == "csv":
-    train_df = pd.read_csv(args.train)
-elif args.train[-3:].lower() == "pkl":
-    train_df = pd.read_pickle(args.train)
-else:
-    raise ValueError("Invalid file format, must be either CSV or PKL.")
-
-if args.test[-3:].lower() == "csv":
-    test_df = pd.read_csv(args.test)
-elif args.test[-3:].lower() == "pkl":
-    test_df = pd.read_pickle(args.test)
-else:
-    raise ValueError("Invalid file format, must be either CSV or PKL.")
-
-print("One-hot encode country, region, and speciesId in training data")
-for column in ["country", "region", "speciesId"]:
-    print(f" - Processing: {column}")
-    train_df = pd.concat(
-        [train_df, pd.get_dummies(train_df[column], prefix=column)],
-        axis=1,
-    )
-    train_df = train_df.drop(columns=[column])
-
-print("One-hot encode country and region in test data")
-for column in ["country", "region"]:
-    print(f" - Processing: {column}")
-    test_df = pd.concat(
-        [test_df, pd.get_dummies(test_df[column], prefix=column)],
-        axis=1,
-    )
-    test_df = test_df.drop(columns=[column])
-
[…]

### 2024-06-06 data-pipeline/5-groupby.py

@@ -1,44 +1,0 @@
-from PIL import Image
-from tqdm import tqdm
-from pathlib import Path
-import pandas as pd
-import argparse as ap
-import os
-
-
-def init_argparse() -> ap.ArgumentParser:
-    parser = ap.ArgumentParser(
-        usage="%(prog)s [OPTION]",
-        description="Convert the satellite images to PNG format.",
-    )
-    parser.add_argument(
-        "--train",
-        type=str,
-        help="The path to the training dataset.",
-        default="output/4-ohe_train.pkl",
-    )
-    parser.add_argument(
-        "--desc",
-        type=str,
-        help="A description of the data set to be used for the file path.",
-        default="grp",
-    )
-    return parser
-
-
-args = init_argparse().parse_args()
-assert os.path.exists(args.train), f"Training data file '{args.train}' not found."
-
-if args.train[-3:].lower() == "csv":
-    train_df = pd.read_csv(args.train)
-elif args.train[-3:].lower() == "pkl":
-    train_df = pd.read_pickle(args.train)
-else:
-    raise ValueError("Invalid file format, must be either CSV or PKL.")
-
-train_df = train_df.groupby("surveyId").max()
-
-print(f"{train_df.shape=}")
-
-train_df.to_pickle(f"output/5-{args.desc}_train.pkl")
-print(f"5-{args.desc}_train.pkl has shape {train_df.shape}")

### 2024-06-06 data-pipeline/6-combine-test.py

@@ -1,74 +1,0 @@
-import pandas as pd
-import argparse as ap
-import os
-
-
-def init_argparse() -> ap.ArgumentParser:
-    parser = ap.ArgumentParser(
-        usage="%(prog)s [OPTION]",
-        description="Convert the satellite images to PNG format.",
-    )
-    parser.add_argument(
-        "--metadata",
-        "-m",
-        type=str,
-        help="The path to the metadata.",
-        default="output/4-ohe_test.pkl",
-    )
-    parser.add_argument(
-        "--additions",
-        "-a",
-        type=str,
-        nargs="+",
-        help="The path to the additional datasets.",
-        default=[
-            "_raw/EnvironmentalRasters/EnvironmentalRasters/Climate/Average 1981-2010/GLC24-PA-test-bioclimatic.csv",
-            # "_raw/EnvironmentalRasters/EnvironmentalRasters/Climate/Monthly/GLC24-PA-test-bioclimatic_monthly.csv",
-            "_raw/EnvironmentalRasters/EnvironmentalRasters/Elevation/GLC24-PA-test-elevation.csv",
-            "_raw/EnvironmentalRasters/EnvironmentalRasters/Human Footprint/GLC24-PA-test-human_footprint.csv",
-            "_raw/EnvironmentalRasters/EnvironmentalRasters/LandCover/GLC24-PA-test-landcover.csv",
-            "_raw/EnvironmentalRasters/EnvironmentalRasters/SoilGrids/GLC24-PA-test-soilgrids.csv",
-            # "_raw/PA-test-landsat_time_series/GLC24-PA-test-landsat_time_series-blue.csv",
-            # "_raw/PA-test-landsat_time_series/GLC24-PA-test-landsat_time_series-green.csv",
-            # "_raw/PA-test-landsat_time_series/GLC24-PA-test-landsat_time_series-nir.csv",
-            # "_raw/PA-test-landsat_time_series/GLC24-PA-test-landsat_time_series-red.csv",
-            # "_raw/PA-test-landsat_time_series/GLC24-PA-test-landsat_time_series-swir1.csv",
-            # "_raw/PA-test-landsat_time_series/GLC24-PA-test-landsat_time_series-swir2.csv",
-        ],
-    )
-    parser.add_argument(
-        "--desc",
-        type=str,
-        help="A description of the data set to be used for the file path.",
-        default="test",
-    )
-    return parser
-
-
[…]

### 2024-06-06 data-pipeline/7-combine-train.py

@@ -1,74 +1,0 @@
-import pandas as pd
-import argparse as ap
-import os
-
-
-def init_argparse() -> ap.ArgumentParser:
-    parser = ap.ArgumentParser(
-        usage="%(prog)s [OPTION]",
-        description="Convert the satellite images to PNG format.",
-    )
-    parser.add_argument(
-        "--metadata",
-        "-m",
-        type=str,
-        help="The path to the metadata.",
-        default="output/5-grp_train.pkl",
-    )
-    parser.add_argument(
-        "--additions",
-        "-a",
-        type=str,
-        nargs="+",
-        help="The path to the additional datasets.",
-        default=[
-            "_raw/EnvironmentalRasters/EnvironmentalRasters/Climate/Average 1981-2010/GLC24-PA-train-bioclimatic.csv",
-            # "_raw/EnvironmentalRasters/EnvironmentalRasters/Climate/Monthly/GLC24-PA-train-bioclimatic_monthly.csv",
-            "_raw/EnvironmentalRasters/EnvironmentalRasters/Elevation/GLC24-PA-train-elevation.csv",
-            "_raw/EnvironmentalRasters/EnvironmentalRasters/Human Footprint/GLC24-PA-train-human_footprint.csv",
-            "_raw/EnvironmentalRasters/EnvironmentalRasters/LandCover/GLC24-PA-train-landcover.csv",
-            "_raw/EnvironmentalRasters/EnvironmentalRasters/SoilGrids/GLC24-PA-train-soilgrids.csv",
-            # "_raw/PA-train-landsat_time_series/GLC24-PA-train-landsat_time_series-blue.csv",
-            # "_raw/PA-train-landsat_time_series/GLC24-PA-train-landsat_time_series-green.csv",
-            # "_raw/PA-train-landsat_time_series/GLC24-PA-train-landsat_time_series-nir.csv",
-            # "_raw/PA-train-landsat_time_series/GLC24-PA-train-landsat_time_series-red.csv",
-            # "_raw/PA-train-landsat_time_series/GLC24-PA-train-landsat_time_series-swir1.csv",
-            # "_raw/PA-train-landsat_time_series/GLC24-PA-train-landsat_time_series-swir2.csv",
-        ],
-    )
-    parser.add_argument(
-        "--desc",
-        type=str,
-        help="A description of the data set to be used for the file path.",
-        default="train",
-    )
-    return parser
-
-
[…]

### 2024-06-06 data-pipeline/8-tidy.py

@@ -1,62 +1,0 @@
-from PIL import Image
-from tqdm import tqdm
-from pathlib import Path
-import pandas as pd
-import numpy as np
-import argparse as ap
-import os
-
-
-def init_argparse() -> ap.ArgumentParser:
-    parser = ap.ArgumentParser(
-        usage="%(prog)s [OPTION]",
-        description="Convert the satellite images to PNG format.",
-    )
-    parser.add_argument(
-        "--train",
-        type=str,
-        help="The path to the training dataset.",
-        default="output/7-combined_train.pkl",
-    )
-    parser.add_argument(
-        "--test",
-        type=str,
-        help="The path to the testing dataset.",
-        default="output/6-combined_test.pkl",
-    )
-    parser.add_argument(
-        "--desc",
-        type=str,
-        help="A description of the data set to be used for the file path.",
-        default="tidy",
-    )
-    return parser
-
-
-args = init_argparse().parse_args()
-assert os.path.exists(args.train), f"Training data file '{args.train}' not found."
-assert os.path.exists(args.test), f"Testing data file '{args.test}' not found."
-
-if args.train[-3:].lower() == "csv":
-    train_df = pd.read_csv(args.train)
-elif args.train[-3:].lower() == "pkl":
-    train_df = pd.read_pickle(args.train)
-else:
-    raise ValueError("Invalid file format, must be either CSV or PKL.")
-
-if args.test[-3:].lower() == "csv":
-    test_df = pd.read_csv(args.test)
-elif args.test[-3:].lower() == "pkl":
-    test_df = pd.read_pickle(args.test)
-else:
-    raise ValueError("Invalid file format, must be either CSV or PKL.")
-
-train_df = train_df.replace([np.inf, -np.inf], np.nan)
-test_df = test_df.replace([np.inf, -np.inf], np.nan)
-
-train_df = train_df.dropna()
-
-train_df.to_pickle(f"output/8-{args.desc}_train.pkl")
-print(f"8-{args.desc}_train.pkl has shape {train_df.shape}")
-test_df.to_pickle(f"output/8-{args.desc}_test.pkl")
-print(f"8-{args.desc}_test.pkl has shape {test_df.shape}")

### 2024-06-06 data-pipeline/9-data-internal.py

@@ -1,74 +1,0 @@
-# Data Processor
-#
-# Desired output
-# - Full train dataset - for training model for Kaggle submission
-# - Full test dataset - for Kaggle submission
-# - Internal train dataset - 80% of full train dataset for internal banchmarking
-# - Internal test dataset - 20% of full train dataset for internal benchmarking
-# - All tabular data in one file
-
-import pandas as pd
-import shutil
-from tqdm import tqdm
-from pathlib import Path
-import argparse as ap
-import os
-
-
-def init_argparse() -> ap.ArgumentParser:
-    parser = ap.ArgumentParser(
-        usage="%(prog)s [OPTION]",
-        description="Convert the satellite images to PNG format.",
-    )
-    parser.add_argument(
-        "--train",
-        type=str,
-        help="The path to the training dataset.",
-        default="output/8-tidy_train.pkl",
-    )
-    parser.add_argument(
-        "--desc",
-        type=str,
-        help="A description of the data set to be used for the file path.",
-        default="int-split",
-    )
-    return parser
-
-
-args = init_argparse().parse_args()
-assert os.path.exists(args.train), f"Training data file '{args.train}' not found."
-
-if args.train[-3:].lower() == "csv":
-    train_df = pd.read_csv(args.train)
-elif args.train[-3:].lower() == "pkl":
-    train_df = pd.read_pickle(args.train)
-else:
-    raise ValueError("Invalid file format, must be either CSV or PKL.")
-
-
-internal_train_df = train_df.sample(frac=0.8, random_state=113)
-internal_test_df = train_df.drop(internal_train_df.index)
-
-print(f"{internal_train_df.shape=}")
-print(f"{internal_test_df.shape=}")
-
-
-internal_train_df.to_pickle(f"output/9-{args.desc}-train.pkl")
-print(f"9-{args.desc}-train.pkl has shape {internal_train_df.shape}")
-
-Path("output/9-internal-train-images").mkdir(parents=True, exist_ok=True)
-for survey_id in tqdm(internal_train_df.surveyId):
-    shutil.copy(
-        f"output/2-all_images/train/{survey_id}.png",
-        f"output/9-internal-train-images/{survey_id}.png",
-    )
-
-Path("output/9-internal-test-images").mkdir(parents=True, exist_ok=True)
-for survey_id in tqdm(internal_train_df.surveyId):
-    shutil.copy(
[…]

### 2024-06-06 data-pipeline/run_me.sh

@@ -1,12 +1,0 @@
-#!/usr/bin/env bash
-
-python 1-top-n.py
-python 2-img-to-png.py
-python 3-metadata-tidy.py
-python 4-onehotencode.py
-python 5-groupby.py
-python 6-combine-test.py
-python 7-combine-train.py
-python 8-tidy.py
-python 9-data-internal.py
-python 10-finalise.py

### 2024-06-06 pipeline2/__init__.py
finished new postprocessor, tidied up a bit more

@@ -3,4 +3,6 @@ from .preprocessing.pcamanager import GeoLifePCA
 
 from .model_management.xgb_manager import GeoLifeXGB
 
-from .postprocessing.postprocessor import GeoLifePostprocessor
+from .postprocessing.postprocessor import (
+    GeoLifePostprocessor, GeoLifeOutputLoader
+)

### 2024-06-06 pipeline2/postprocessing/postprocessor.py

@@ -1,17 +1,47 @@
+import numpy as np
+import pickle as pk
 import pandas as pd
 from datetime import datetime 
 from .applycounts import apply_counts
 
+class GeoLifeOutputLoader():
+
+    def __init__(self, species, surveys):
+        self.species = species 
+        self.surveys = surveys 
+
+    def load_outputs(self, path_list:list):
+        outputs = []
+        for path in path_list:
+            loaded_data = (pd.read_pickle(path))
+            if type(loaded_data) == np.ndarray:
+                loaded_data = pd.DataFrame(
+                    loaded_data,
+                    columns=self.species,
+                    index=self.surveys
+                )
+            else:
+                loaded_data = loaded_data[self.species]
+                loaded_data = loaded_data.set_index(self.surveys)
+            outputs.append(loaded_data)
+        return outputs
+    
+    def load_counts(self, path):
+        return pk.load(open(path, "rb"))
+
+
 class GeoLifePostprocessor():
 
     def __init__(self, predictions, weights, counts):
+        self.index = predictions[0].index
         self.predictions = None 
         for prediction, weight in zip(predictions, weights):
-            prediction = prediction.to_numpy()
+            pred = prediction.to_numpy()
             if self.predictions is None: 
-                self.predictions = prediction * weight 
+                self.predictions = pred * weight 
             else:
-                self.predictions += prediction * weight 
+                self.predictions += pred * weight 
+        pd.DataFrame(self.predictions).to_csv("test-new.csv", index=False)
         self.predictions = apply_counts(self.predictions, counts)
         self.predictions = pd.DataFrame(
             self.predictions, 
@@ -26,6 +56,7 @@ class GeoLifePostprocessor():
             f"submissions/{filename}", 
             index=False
         )
+        print(f"Saved to: submissions/{filename}")
     
     def __format_submission(self):
         output_list = []
@@ -36,6 +67,6 @@ class GeoLifePostprocessor():
             output_list.append(row_string)
 
         return pd.DataFrame({
-            "surveyId"      :   self.predictions.index,
+            "surveyId"      :   self.index,
             "predictions"   :   output_list
         })

### 2024-06-06 pipeline2/preprocessing/landsat.py

@@ -2,8 +2,6 @@ import numpy as np
 import pandas as pd
 import pickle as pk
 
-TRAIN_IDS = pk.load(open("data/processed/train.pkl", "rb")).index
-TRAIN_IDS = TRAIN_IDS.to_list()
 COLORS = ["red", "green", "blue", "nir", "swir1", "swir2"]
 
 def get_path(is_train, color):
@@ -12,14 +10,15 @@ def get_path(is_train, color):
     path += f"-landsat_time_series-{color}.csv"
     return path 
 
-def pickle_landsat():
+def pickle_landsat(metadata_train):
+    train_ids = metadata_train.index.to_list()
     for truth in [True, False,]:
         image_data = None 
         for c_idx, color in enumerate(COLORS):
             path = get_path(truth, color)
             data = pd.read_csv(path)
             if image_data is None:
-                axis_0 = len(TRAIN_IDS) if truth else data.shape[0]
+                axis_0 = len(train_ids) if truth else data.shape[0]
                 image_data = np.zeros((axis_0, 60, 6), np.uint8)
             cols = data.columns.to_list()
             cols.reverse()
@@ -31,7 +30,7 @@ def pickle_landsat():
             data.reset_index()
             idx = 0
             for _, row in data.iterrows():
-                if truth and row["surveyId"] not in TRAIN_IDS: 
+                if truth and row["surveyId"] not in train_ids: 
                     continue
                 
                 for modifier in range(0, 15):
@@ -45,7 +44,3 @@ def pickle_landsat():
         save_path = f"data/processed/{'train' if truth else 'test'}"
         save_path += "_ls.pkl"
         pk.dump(image_data, open(save_path, "wb"))
-
-
-if __name__ == "__main__":
-    pickle_landsat()

### 2024-06-06 helper/__init__.py
finished new postprocessor, tidied up a bit more

@@ -1,7 +1,0 @@
-""" import helper
-"""
-# from .verifytest import verify_test
-# from .verifysubmission import verify_submission
-# from .images_to_np import load_images
-# from .tifnormalizer import tif_helper
-from .top_n_regression import get_top_n

### 2024-06-06 helper/conda-installs.sh

@@ -1,32 +1,0 @@
-ba_flag=''
-tf_flag=''
-pt_flag=''
-
-while getopts 'btp' flag; do
-    case "${flag}" in
-        b)  ba_flag="true" ;;
-        t)  tf_flag="true" ;;
-        p)  pt_flag="true" ;;
-        *)  echo "exiting" && exit 1 ;;
-    esac
-done
-
-if [ $ba_flag -eq "true" ]; then
-    echo ":: [INSTALL] Base packages" ;
-    conda activate geolife ;
-    conda install anaconda::numpy anaconda::matplotlib anaconda::ipykernel conda-forge::pandas conda-forge::tqdm conda-forge::opencv
-fi
-
-if [ $tf_flag -eq "true" ]; then
-    echo ":: [INSTALL] Tensorflow" ;
-    conda activate geolife ;
-    conda install conda-forge::tensorflow ;
-fi
-
-if [ $pt_flag -eq "true" ]; then
-    echo ":: [INSTALL] PyTorch" ;
-    conda activate geolife ;
-    conda install pytorch torchvision torchaudio pytorch-cuda=11.8 -c pytorch -c nvidia ;
-fi
-
-

### 2024-06-06 helper/images_to_np.py

@@ -1,39 +1,0 @@
-""" from helper import load_images
-"""
-
-import os
-import cv2
-import numpy as np
-from tqdm import tqdm
-
-def load_images(labels:list, is_train:bool = True):
-    """Imports four channel png images from either the train or test data set
-    and returns them as a numpy array ordered by the id list provided.
-
-    args:
-        labels          :   List of the surveyId labels corresponding to the 
-                            desired images
-        is_train        :   True if images are from the training data, False if
-                            from the testing data
-
-    return:
-        np.array            An array of int8 values containg the loaded images
-    """
-    if not os.path.isdir("data-pipeline"):
-        raise Exception(":: [ERROR] this is designed to be run from the root " \
-                        + "directory, however cannot find \'data-pipeline\'")
-    images = np.zeros((len(labels), 128, 128, 4), dtype=np.uint8)
-    image_path = os.path.join(
-        "data-pipeline/output/2-all_images",
-        "train" if is_train else "test"
-    )
-    for idx, img_id in enumerate(tqdm(labels)):
-        path = os.path.join(image_path, f"{img_id}.png")
-        try:
-            images[idx] = cv2.imread(path, cv2.IMREAD_UNCHANGED)
-        except:
-            message = f":: [ERROR] image with ID {img_id} not found at "
-            message += f"\'{path}\'"
-            raise Exception(message)
-    return images
-    

### 2024-06-06 helper/tifnormalizer.py

@@ -1,26 +1,0 @@
-""" from helper import tif_helper
-"""
-import cv2 
-import numpy as np
-
-def tif_helper(filepath:str):
-    image = cv2.imread(filepath, cv2.IMREAD_UNCHANGED)
-    min_val = image.min()
-
-    print(min_val)
-    
-    if min_val > 0:
-        return image 
-    
-    previous_value = (image * (image != min_val)).mean()
-    dims = image.shape
-
-    for ii in range(dims[0]):
-        for jj in range(dims[1]):
-            if image[ii, jj] == min_val:
-                image[ii, jj] = previous_value
-            previous_value = image[ii, jj]
-
-    # if filepath
-
-    return image

### 2024-06-06 helper/top_n_regression.py

@@ -1,15 +1,0 @@
-import numpy as np
-
-def get_top_n(data, n):
-    topn = np.zeros(data.shape, dtype=bool)
-    if type(n) is not int:
-        if len(n) != data.shape[0]:
-            print(":: [ERROR] n must either be an integer or an array of equal"\
-                  " length to the input data")
-            return None 
-        for idx in range(data.shape[0]):
-            true_n = n if type(n) is int else n[idx]
-            for col in np.argpartition(data[idx, :], -true_n)[-true_n:]:
-                topn[idx, col] = 1
-    return topn
-

### 2024-06-06 helper/verifysubmission.py

@@ -1,45 +1,0 @@
-"""Verifies that the contents of a submission file appears to be correct
-"""
-
-import os
-import pandas as pd
-
-def verify_submission(filename:str,
-                      raw_path:str = "data-pipeline/_raw") -> bool:
-    """Checks that the shape and surveyId values of a submission CSV file are
-    correct
-
-    args:
-        filename        :   The name of the submission file to be checked
-        raw_path        :   The directory in which the sample submission is
-                            located
-
-    return:
-        bool            :   True if the file appears correct, otherwise False
-    """
-    sample_submission = pd.read_csv(
-        os.path.join(raw_path, "GLC24_SAMPLE_SUBMISSION.csv")
-    )
-    generated_submission = pd.read_csv(os.path.join("submissions", filename))
-
-    if sample_submission.shape != generated_submission.shape:
-        print(":: Shape of dataframes is not equal.")
-        return False
-
-    correct_ids = True
-    for sample_id in sample_submission["surveyId"]:
-        if sample_id not in generated_submission["surveyId"].to_list():
-            print(f":: ID {sample_id} in sample but not predictions")
-            correct_ids = True
-    for prediction_id in generated_submission["surveyId"]:
-        if prediction_id not in sample_submission["surveyId"].to_list():
-            print(f":: ID {prediction_id} in predictions but not in sample")
-            correct_ids = False
-
-    true_cols = ["surveyId", "predictions"]
-    gen_cols  = list(generated_submission.columns)
-    correct_cols = true_cols[0] == gen_cols[0] and true_cols[1] == gen_cols[1]
-    if not correct_cols:
-        print(f":: Column names should be {true_cols}, not {gen_cols}")
-
-    return correct_ids and correct_cols

### 2024-06-06 helper/verifytest.py

@@ -1,29 +1,0 @@
-"""Verifies that the processed test data is valid.
-"""
-
-import os
-import pandas as pd
-
-def verify_test(dataframe:pd.DataFrame,
-                dir_path:str = "data-pipeline/_raw") -> bool:
-    """Verify that the surveyId values in the processed data match those in the
-    raw test data.
-    """
-    processed_cols = dataframe["surveyId"] if "surveyId" in dataframe.columns \
-                                         else dataframe.index
-    original_cols = pd.read_csv(
-        os.path.join(dir_path, "GLC24_PA_metadata_test.csv")
-    )["surveyId"]
-
-    only_in_proc = list(set(processed_cols) - set(original_cols))
-    only_in_orig = list(set(original_cols) - set(processed_cols))
-
-    message = ""
-    if len(only_in_proc) > 0:
-        message += f":: {len(only_in_proc)} IDs only in the processed data.\n"
-    if len(only_in_orig) > 0:
-        message += f":: {len(only_in_orig)} IDs only in the raw data.\n"
-    if len(message) > 0:
-        print(message)
-
-    return len(only_in_orig) + len(only_in_proc) == 0 and correct_cols

### 2024-06-06 pipeline/getdata/__init__.py

@@ -1,1 +1,0 @@
-from .setupdirs import DirectoryManagement 

### 2024-06-06 pipeline/getdata/downloaddata.sh

@@ -1,8 +1,0 @@
-if [ -z "$(ls -A data/_raw)" ]; then
-    echo ":: Downloading raw data to data/_raw"
-    cd data/_raw
-    kaggle competitions download -c geolifeclef-2024
-    cd -
-else
-    echo ":: data/_raw is not empty. Aborting download."
-fi

### 2024-06-06 pipeline/getdata/setupdirs.py

@@ -1,41 +1,0 @@
-""" from pipeline.getdata import DirectoryManagement
-"""
-import os
-import subprocess
-
-class DirectoryManagement():
-    """ Ensures that the correct directory structure is in place before using
-    a bash script to download the raw data from kaggle
-    """
-
-    dir_path = "data/"
-    raw_path = "data/_raw/"
-    processed_path = "data/processed/"
-
-    def __init__(self, **kwargs):
-        """ Interprets kwargs and calls relevant functions
-        
-        kwargs:
-            dirs        :   False to not check and create directories
-            download    :   False to not download data
-        """
-        if "dirs" not in kwargs or not kwargs.pop("dirs"):
-            self.check_make_dirs()
-        if "download" not in kwargs or not kwargs.pop("downloads"):
-            self.download_data()
-
-    def check_make_dirs(self):
-        """ Checks if directories exist, and creates them if not.
-        """
-        print(":: Setting up dirs:\n\tdata/_raw/\n\t    /processed/")
-        if not os.path.exists(self.dir_path):
-            os.mkdir(self.dir_path)
-        if not os.path.exists(self.raw_path):
-            os.mkdir(self.raw_path)
-        if not os.path.exists(self.processed_path):
-            os.mkdir(self.processed_path)
-
-    def download_data(self):
-        """ Runs a bash script to download data from kaggle
-        """
-        subprocess.run(['bash', 'pipeline/getdata/downloaddata.sh'])

### 2024-06-06 pipeline/modelprocessing/__init__.py

@@ -1,1 +1,0 @@
-from .modelproc import save_model, load_model

### 2024-06-06 pipeline/modelprocessing/modelproc.py

@@ -1,29 +1,0 @@
-import os
-import pickle
-import datetime
-
-def save_model(model, model_type: str, note: str = None, pred_cols = None):
-    now_string = datetime.datetime.now().strftime("%H-%M_%d-%m")
-    build_dirs(model_type, now_string)
-    note = 'model' if note is None else note
-    full_path = os.path.join("models", model_type, now_string, f"{note}.pkl")
-    pickle.dump(model, open(full_path, "wb"))
-    print(f":: Saved model to {full_path}")
-    if pred_cols is not None:
-        pickle.dump(pred_cols, open(full_path[:-4] + '_cols.pkl', "wb"))
-
-def build_dirs(model_type, timestamp):
-    for dir in ["models/", f"models/{model_type}", 
-                f"models/{model_type}/{timestamp}"]:
-        if not os.path.exists(dir):
-            print(f":: Creating directory {dir}")
-            os.mkdir(dir)
-
-def load_model(model_type: str, model_name: str, timestamp: str = None):
-    if len(model_name) < 4 or model_name[-4:] != '.pkl':
-        model_name += '.pkl'
-    if timestamp is not None:
-        path = os.path.join("models", model_type, timestamp, model_name)
-    else:
-        path = os.path.koin("models", model_type, model_name)
-    return pickle.load(open(path, 'rb'))

### 2024-06-06 pipeline/postprocessing/__init__.py

@@ -1,3 +1,0 @@
-from .postprocessor import Postprocessor
-from .exceptions import ShapeError
-from .exceptions import MissingDataError

### 2024-06-06 pipeline/postprocessing/exceptions.py

@@ -1,41 +1,0 @@
-"""Exceptions and errors for the postprocessor
-"""
-
-class ShapeError(Exception):
-    """An error for when the predictions and labels do not have the same number 
-    of columns
-    """
-    message = ":: [ERROR] Predictions and labels must have an equal number of "\
-              "rows. If this is for a submission run again with uploadable=True"
-
-    def __init__(self) -> None:
-        """Passes the message directly to the Exception class
-        """
-        super().__init__(self.message)
-
-
-class MissingDataError(Exception):
-    """An error when insufficient data has been passed to construct the 
-    predictions dataframe
-    """
-
-    def __init__(self, columns, indices) -> None:
-        """Passes the message directly to the Exception class
-        """
-        super().__init__(self.__write_message(columns, indices))
-
-    def __write_message(self, columns, indices):
-        """Constructs the error message
-
-        args:
-            columns     :   The data passed in for the column names
-            indices     :   The data passed in for the indice values
-        """
-        col_section = ":: Missing pred_cols argument.\n-\tPlease pass a list " \
-                    + "of column values. These could be taken from the "       \
-                    + "pd.DataFrame.columns value of the training labels.\n"
-        idx_section = ":: Missing pred_indices argument.\n-\tPlease pass a "   \
-                    + "list of index values. These could be taken from the "   \
-                    + "pd.DataFrame.index value of the test input data\n"
-        return col_section if not isinstance(columns, list) else "" +          \
-               idx_section if not isinstance(indices, list) else ""

### 2024-06-06 pipeline/postprocessing/postprocessor.py

@@ -1,194 +1,0 @@
-"""
-# from postprocessing.postprocessor import Postprocessor
-
-Postprocessor for prediction data. This is designed to be called from notebooks
-in the root directory.
-
-    - Converts data to dataframe of booleans with given index and column names
-    - Performs f1 score on the data
-    - Formats submission and saves it to the 'submissions' directory
-"""
-import os
-import re
-from datetime import datetime
-
-import numpy as np
-import pandas as pd
-from sklearn.metrics import f1_score
-
-from .exceptions import MissingDataError, ShapeError
-
-class Postprocessor():
-    """Performs post processing operations on the raw prediction data generated
-    by a model. Also provides an f1 score when labels are available.
-    """
-    f1_score = None   # Placeholder for the f1 score
-
-    def __init__(self, predictions, labels = None, pred_cols = None,
-                 pred_indices = None, **kwargs) -> None:
-        """
-        Converts the predictions to a DataFrame and parses kwargs
-
-        args:
-            predictions     :   Either a numpy array or a pandas DataFrame of 
-                                predicted values
-            labels          :   A pandas DataFrame of the labels corresponding 
-                                to the predictions
-            pred_cols       :   [Not needed if predictions is a DataFrame] 
-                                A list of the column names for the predictions
-            pred_indices    :   [Not needed if predictions is a DataFrame] 
-                                A list of the indice values for the predictions
-
-        kwargs:
-            f1              :   True to perform a f1 score immediately
-            save            :   True to save to default submission path                    
-            model_type      :   The type of model for the saved filename
-            uploadable      :   True if the predictions are from the kaggle
-                                test data
-        """
-        self.uploadable = (
-            "uploadable" in kwargs and kwargs.pop("uploadable")
-        )
-
-        if not self.uploadable and predictions.shape[0] != labels.shape[0]:
-            raise ShapeError()
-        self.labels = labels
-        self.predictions = predictions
-        if isinstance(predictions, np.ndarray):
-            self.__preds_to_df(list(pred_cols), list(pred_indices))
-
-        self.__normalise_columns()
-
-        self.__parse_kwargs(**kwargs)
-
-
-    def get_f1_score(self, **kwargs) -> np.float64:
-        """Formats the predictions dataframe so that it matches the labels 
-        dataframe in structure then calculates and prints the f1 score to three
-        decimal places. Stores the untruncated value as a member variable.
-        """
[…]

### 2024-06-06 pipeline/preprocess/__init__.py

@@ -1,4 +1,0 @@
-from .landsat_combiner import pickle_landsat
-from .preprocessor import DataCollector
-from .dataloader import load_x_y, load_with_split, match_test_columns
-from .pca import PCACreator, PCAApplier

### 2024-06-06 pipeline/preprocess/dataloader.py

@@ -1,123 +1,0 @@
-""" 
-from preprocess.dataloader import load_x_y
-from preprocess.dataloader import load_with_split
-"""
-
-import numpy as np
-import pandas as pd
-from sklearn.model_selection import train_test_split
-
-
-def _handle_drops(data, drop_type):
-    """Helper function.
-    Handles the "drop" key word argument
-    """
-    if "soil" in drop_type or "def" in drop_type:
-        data = data.drop(columns=[
-            "Soilgrid-bdod",
-            "Soilgrid-cec",
-            "Soilgrid-cfvo",
-            "Soilgrid-clay",
-            "Soilgrid-nitrogen",
-            "Soilgrid-phh2o",
-            "Soilgrid-sand",
-            "Soilgrid-silt",
-            "Soilgrid-soc",
-        ])
-    if "small" in drop_type or "def" in drop_type: 
-        data = data.dropna(subset=[
-            "Elevation",
-        ])
-    if "safe" in drop_type or "def" in drop_type:
-        data["geoUncertaintyInM"] = data["geoUncertaintyInM"].replace(
-            np.nan, 
-            data["geoUncertaintyInM"].median(),
-        )
-    if "human" in drop_type or "def" in drop_type:
-        data = data.drop(columns=[
-            "HumanFootprint-NavWater1994",
-            "HumanFootprint-NavWater2009",
-            "HumanFootprint-Roads",
-            "HumanFootprint-HFP1993",
-            "HumanFootprint-HFP2009",
-        ])
-    return data
-
-def _handle_area(data, area_type):
-    """Helper function.
-    Handles the "area" key word argument
-    """
-    if area_type == "dropcol":
-        return data.drop(columns="areaInM2")
-    if area_type == "droprow":
-        return data.dropna(subset=["areaInM2"])
-    if area_type == "median":
-        data["areaInM2"] = data["areaInM2"].fillna(data["areaInM2"].median())
-        return data
-    if area_type == "mean":
-        data["areaInM2"] = data["areaInM2"].fillna(data["areaInM2"].mean())
-        return data
-    return data
-
-def _handle_filtering(data, filter_method):
-    if not filter_method:
-        return data 
-    filter_mask = [col for col in data if col.startswith('Bio-') and           \
-                   '06_2018' not in col and '12_2018' not in col]
-    return data.drop(columns=filter_mask)
-
-def load_x_y(is_train: bool, **kwargs):
[…]

### 2024-06-06 pipeline/preprocess/landsat_combiner.py

@@ -1,54 +1,0 @@
-import numpy as np
-import pandas as pd
-import pickle as pk
-
-from .pca import PCAApplier 
-
-TRAIN_IDS = PCAApplier("train", method="full", from_year=2010).data.index
-TRAIN_IDS = TRAIN_IDS.to_list()
-COLORS = ["red", "green", "blue", "nir", "swir1", "swir2"]
-
-def get_path(is_train, color):
-    is_train = "train" if is_train else "test"
-    path = f"data/_raw/PA-{is_train}-landsat_time_series/GLC24-PA-{is_train}"
-    path += f"-landsat_time_series-{color}.csv"
-    return path 
-
-def pickle_landsat():
-    for truth in [True, False,]:
-        image_data = None 
-        for c_idx, color in enumerate(COLORS):
-            print(f":: [{'TRAIN' if truth else 'TEST'}] -- {color}")
-            path = get_path(truth, color)
-            data = pd.read_csv(path)
-            if image_data is None:
-                axis_0 = len(TRAIN_IDS) if truth else data.shape[0]
-                image_data = np.zeros((axis_0, 60, 6), np.uint8)
-            cols = data.columns.to_list()
-            cols.reverse()
-            data = data[cols]
-            data["from"] = data.apply(
-                lambda x: str(int(x.first_valid_index()[:4]) - 15) + "_1",
-                axis=1
-            )
-            data.reset_index()
-            idx = 0
-            for _, row in data.iterrows():
-                if truth and row["surveyId"] not in TRAIN_IDS: 
-                    continue
-                
-                for modifier in range(0, 15):
-                    for quarter in range(1, 5):
-                        image_data[
-                            idx, 4 * modifier + quarter - 1, c_idx
-                        ] = np.uint8(
-                            row[f"{int(row['from'][:4]) + modifier}_{quarter}"]
-                        )
-                idx += 1
-        save_path = f"data/processed/full_{'train' if truth else 'test'}"
-        save_path += "_landsat.pkl"
-        pk.dump(image_data, open(save_path, "wb"))
-
-
-if __name__ == "__main__":
-    pickle_landsat()

### 2024-06-06 pipeline/preprocess/pca.py

@@ -1,228 +1,0 @@
-""" from pipeline.preprocess import PCACreator, PCAApplier
-"""
-import os
-import pickle as pk
-import pandas as pd
-from warnings import simplefilter
-from numpy import float32, float64, inf, nan
-from sklearn.preprocessing import scale
-
-# Hide irrelevant "performance wanrning" from pandas library 
-simplefilter(action="ignore", category=pd.errors.PerformanceWarning)
-
-class PCAGeneric():
-    """ Generic class providing methods and variables to PCACreator and 
-    PCAApplier classes
-    """
-
-    rw_path = "models/pca"
-
-    high_cov_hf_cols = [
-        'HumanFootprint-Built1994', 
-        'HumanFootprint-Built2009', 
-        'HumanFootprint-croplands1992', 
-        'HumanFootprint-croplands2005', 
-        'HumanFootprint-Lights1994', 
-        'HumanFootprint-Lights2009', 
-        'HumanFootprint-Pasture1993', 
-        'HumanFootprint-Pasture2009', 
-        'HumanFootprint-Popdensity1990', 
-        'HumanFootprint-Popdensity2010', 
-        'HumanFootprint-Railways',
-    ]
-    drop_cols = [
-        "HumanFootprint-HFP1993", 
-        "HumanFootprint-NavWater1994", 
-        "HumanFootprint-Roads",
-        "Latvia",
-    ]
-    tas_prefixes = [
-        "Bio-tas_",
-        "Bio-tasmax_",
-        "Bio-tasmin_",
-    ]
-    pr_prefix = [
-        "Bio-pr_",
-    ]
-
-    def __init__(self, path, **kwargs):
-        path = os.path.join("data/processed", path + "_full.pkl")
-        self.data = pd.read_pickle(path)
-        if "from_year" in kwargs:
-            self.from_year = kwargs.pop("from_year")
-            self.__drop_years()
-        
-
-    def apply_human_footprint(self):
-        """Applies the PCA model for human footprint, combining several 
-        columns and deleting the originals
-        """
-        pca = self._load("pca_hf")
-        self.data["HumanFootprintGeneric"] = pca.transform(
-            self.data[self.high_cov_hf_cols]
-        )
-        self.data.drop(columns=self.high_cov_hf_cols, inplace=True)
-
-    def apply_rainfalls(self):
-        """Applies the PCA model for rainfall and temperature, combining several 
-        columns and deleting the originals
-        """
-        self._process_rainfall(
[…]

### 2024-06-06 pipeline/preprocess/preprocessor.py

@@ -1,177 +1,0 @@
-""" from pipeline.preprocess import DataCollector
-"""
-
-import os
-import numpy as np
-import pandas as pd
-
-class DataCollector():
-    """ Collects the raw data and combines it into a set of single master 
-    dataframes
-    """
-
-    def __init__(self, base_path: str = None, raster_path: str = None,
-                 drop_list: list = None, reassign_list: list = None) -> None:
-        """ Set default values if no custom values passed in
-        """
-        self._base_path = "data/_raw" if base_path is None else base_path
-        self._raster_path = "EnvironmentalRasters/EnvironmentalRasters" if     \
-                            raster_path is None else raster_path
-        self._drop_list = [         # Occurences after group by survey:
-            "PANNONIAN",            # 248 occurences 
-            "STEPPIC",              # 48 occurences 
-            "BLACK SEA",            # 8 occurences
-            "Hungary",              # 29 occurences
-            "Ireland",              # 12 occurences
-            "Monaco",               # 7 occurences
-            "The former Yugoslav Republic of Macedonia", # 6
-            "Portugal",             # 6 occurences
-            "Norway",               # 5 occurences
-            "Andorra",              # 4 occurences
-            "Luxembourg",           # 2 occurences
-        ] if drop_list is None else drop_list
-        self._reassign_list = [
-            "Bosnia and Herzegovina",
-            "Bulgaria",
-            "Slovenia",
-            "Greece",
-            "Serbia",
-            "Slovakia",
-            "Croatia",
-        ] if reassign_list is None else reassign_list
-        self.rasters = None
-        self.data = None
-
-    def process_and_save_data(self) -> None:
-        """ Loads in and processes the data for both test and train datasets 
-        """
-        for is_train in [True, False]:
-            purpose = "train" if is_train else "test"
-            filename = purpose + "_full.pkl"
-            filepath = os.path.join("data", "processed", filename)
-            print(f":: [{purpose.upper()}] Loading in Environmental Rasters")
-            self.combine_rasters(is_train=is_train)
-            print(f":: [{purpose.upper()}] Loading in core data")
-            self.get_core_data(is_train=is_train)
-            print(f":: [{purpose.upper()}] One hot encoding data")
-            self.encode_data(is_train=is_train)
-            print(f":: [{purpose.upper()}] Saving data to \'{filepath}\'")
-            self.data.join(self.rasters).to_pickle(filepath)
-
-
[…]

### 2024-06-06 preprocess_data.py

@@ -1,12 +1,23 @@
+import os
 import pickle as pk
 from pipeline2 import GeoLifePreprocessor, GeoLifePCA
 from pipeline2.preprocessing import pickle_landsat
 
+class SetupException(Exception):
+
+    def __init__(self, *args: object) -> None:
+        super().__init__("Please import the raw data into data/_raw/")
+
 if __name__ == "__main__":
+    if not os.path.exists("data/_raw"):
+        raise SetupException()
+    if not os.path.exists("data/processed"):
+        os.makedirs("data/processed")
+
     train, test = GeoLifePreprocessor().process_data()
     train, test = GeoLifePCA(train, test).process_data()
     pk.dump(train, open("data/processed/train.pkl", "wb"))
     pk.dump(test, open("data/processed/test.pkl", "wb"))
-    train_ls, test_ls = pickle_landsat()
+    train_ls, test_ls = pickle_landsat(train)
     pk.dump(train_ls, open("data/processed/train_ls", "wb"))
     pk.dump(test_ls, open("data/processed/test_ls", "wb"))

### 2024-06-06 dataformm.py
Add mm data processing

@@ -1,0 +1,72 @@
+from torch.utils.data import Dataset
+import torch
+
+import pickle as pk
+import os
+from PIL import Image
+
+# import torchvision.transforms as transforms
+
+
+class TestDataset(Dataset):
+    def __init__(self, transform=None):
+        self.subset = "test"
+        self.pca_data = pk.load(open("data/processed/test.pkl", "rb"))
+        self.landsat_data = pk.load(open("data/processed/test_ls.pkl", "rb"))
+        self.transform = transform
+
+    def __len__(self):
+        return self.pca.shape[0]
+
+    def __getitem__(self, idx):
+        survey_id = self.pca_data.index[idx]
+        image_data = self.process_patch(survey_id)
+        sample = [self.pca_data.iloc[idx].tolist(), image_data, self.landsat_data[idx]]
+        return sample, survey_id
+
+    def process_patch(self, survey_id):
+        rgb_path = f"data/_raw/PA_{self.subset.title()}_SatellitePatches_RGB/pa_{self.subset}_patches_rgb"
+        nir_path = f"data/_raw/PA_{self.subset.title()}_SatellitePatches_NIR/pa_{self.subset}_patches_nir"
+        for d in (str(survey_id)[-2:], str(survey_id)[-4:-2]):
+            rgb_path = os.path.join(rgb_path, d)
+            nir_path = os.path.join(nir_path, d)
+        rgb_path = os.path.join(rgb_path, f"{survey_id}.jpeg")
+        nir_path = os.path.join(nir_path, f"{survey_id}.jpeg")
+        rgb_image = Image.open(rgb_path).convert("RGB")
+        rgb_image = self.transform(rgb_image)
+        rgb_image = rgb_image.unsqueeze(0)
+        nir_image = Image.open(nir_path).convert("L")
+        nir_image = self.transform(nir_image)
+        nir_image = nir_image.unsqueeze(0)
+        image_data = torch.cat([rgb_image, nir_image], dim=1)
+        image_data = torch.squeeze(image_data)
+        return image_data
+
+
+class TrainDataset(TestDataset):
+    def __init__(self, transform=None):
+        self.subset = "train"
+        self.pca_data = pk.load(open("data/processed/train.pkl", "rb"))
+        self.landsat_data = pk.load(open("data/processed/train_ls.pkl", "rb"))
+        self.transform = transform
+        label_cols = [col for col in self.pca_data.columns if "speciesId_" in col]
[…]

### 2024-06-23 pipeline/preprocessing/preprocessor.py
Minor updates to preprocessor

@@ -40,14 +40,17 @@ class GeoLifePreprocessor():
             "EnvironmentalRasters"
         )
 
-    def process_data(self) -> None: 
+    def process_data(self, species_drop: int=100) -> None: 
         """ Loads in and processes the data for both test and train datasets
         """
         train = None 
         test = None
         for purpose in ["train", "test"]:
             self._combine_rasters(is_train=(purpose == "train"))
-            self._get_metadata(is_train=(purpose == "train"))
+            self._get_metadata(
+                is_train=(purpose == "train"), 
+                species_drop=species_drop
+            )
             self._encode_data(is_train=(purpose == "train"))
             self._data.replace([np.inf, -np.inf], np.nan, inplace=True)
             self._data = self._data.fillna(self._data.median())

### 2024-06-23 preprocess_data.py

@@ -1,7 +1,8 @@
 import os
+import sys
 import pickle as pk
-from pipeline2 import GeoLifePreprocessor, GeoLifePCA
-from pipeline2.preprocessing import pickle_landsat
+from pipeline import GeoLifePreprocessor, GeoLifePCA
+from pipeline.preprocessing import pickle_landsat
 
 class SetupException(Exception):
 
@@ -9,15 +10,26 @@ class SetupException(Exception):
         super().__init__("Please import the raw data into data/_raw/")
 
 if __name__ == "__main__":
+
+    species_drop = 100
+
+    args = sys.argv
+    if '--species' in args and len(args) > args.index('--species') + 1:
+        species = args[args.index('--species') + 1]
+        if species.isdigit() and int(species) > 0 and int(species) < 1000:
+            species_drop = int(species)
+    print(f"Dropping species below {species_drop} occurrences")
+
     if not os.path.exists("data/_raw"):
         raise SetupException()
     if not os.path.exists("data/processed"):
         os.makedirs("data/processed")
 
-    train, test = GeoLifePreprocessor().process_data()
+    train, test = GeoLifePreprocessor().process_data(species_drop=species_drop)
     train, test = GeoLifePCA(train, test).process_data()
     pk.dump(train, open("data/processed/train.pkl", "wb"))
     pk.dump(test, open("data/processed/test.pkl", "wb"))
-    train_ls, test_ls = pickle_landsat(train)
-    pk.dump(train_ls, open("data/processed/train_ls", "wb"))
-    pk.dump(test_ls, open("data/processed/test_ls", "wb"))
+    # train_ls, test_ls = pickle_landsat(train)
+    pickle_landsat(train)
+    # pk.dump(train_ls, open("data/processed/train_ls", "wb"))
+    # pk.dump(test_ls, open("data/processed/test_ls", "wb"))

### 2024-06-23 paper/sections/00-title-abstract.tex
Add introduction first draft
Minor typo in readme

@@
 %% This command is for the conference information            %%%%%%% TODO !!! UPDATE THIS %%%%%%
-\conference{Woodstock'22: Symposium on the irreproducible science,
-  June 07--11, 2022, Woodstock, NY}
+\conference{CLEF 2024: Conference and Labs of the Evaluation Forum, September 9-12, 2024, Grenoble, France}
 
@@
 ]
+\fnmark[1]
 \address[1]{University of Groningen, Broerstraat 5, 9712 CP Groningen, Netherlands}
@@
 ]
+\fnmark[1]
 
@@
 \begin{keywords}
-  LaTeX class \sep
-  paper template \sep
-  paper formatting \sep
+  Multi-Label classification\sep
+  Principal component analysis\sep
+  ResNet\sep
+  Vision Transformers\sep
+  XGBoost\sep
+  GeoLifeCLEF 2024\sep
   CEUR-WS

### 2024-06-23 paper/sections/04-method.tex
Minor typo in readme

--- 2024-05-29 paper/sections/04-method.tex
@@
-\clearpage  % TODO comment this out when finished
-
 \section{Method}
 
-In order to manage the multilabel classification required for this research we implemented an ensemble approach.
-This section details the individual architectures and the methods used to combine them.
+In order to manage the multilabel classification required for this research we implemented three main approaches as well as ensemble combinations of each approach.
+This section details the data preprocessing as well as the individual architectures and the methods used to combine them.
+
+\subsection{Preprocessing}
+
+The Prsence-Absence Metadata, Environmental rasters and Time-Series Satellite data were preprocessed.
+
+\begin{figure}\label{fig:preproc}
+    \begin{center}
+        \input{figs/pca-diagram.tex}
+        \caption{Preprocessing the data for use in the count prediction and XGBoost pseudo-probability models}
+    \end{center}
+\end{figure}
+
+\subsubsection{Dimensionality Reduction with Principal Component Analysis}
+Principal Component Analysis (PCA) is a method for reducing dimensionality~\cite{pcapaper}. 
+It functions by projecting the high dimensional data onto the direction of maximum variance, thus retaining key features and reducing noise.
+
+We applied PCA to much of the Time Series data within the Environmental Rasters. Each month in a given class
+
+\subsubsection{Missing and infinite values}
+
+\subsubsection{Low-Value Data Removal}
+
+% The raw data included some missing, or infinite values that needed to be processed prior to model fitting.
+When a column was both deemed important to model fitting and contained such values, these values were replaced by the mean so as to avoid excessive influence from outliers without overly effecting the shape of the data~\cite{missingmedian}.
+% 
+% \textbf{TODO: Further explanation when models finalized}
+
+% The Metadata and Bioclimatic Rasters were preprocessed for use in the count and XGBoost models in order to both reduce dimensionality and remove potential noise.
 
@@
 then weighted and combined before the maximum arguments were selected.
-
-\subsubsection{Principal Component Analysis}
-Principal Component Analysis (PCA) is a method for reducing dimensionality~\cite{pcapaper}.
-It functions by projecting the high dimensional data onto the direction of maximum variance, thus retaining key features and reducing noise.
 
@@
 
-\subsubsection{Preprocessing}
-
-\begin{figure}\label{fig:preproc}
-    \begin{center}
-        \input{figs/pca-diagram.tex}
-        \caption{Preprocessing the data for use in the count prediction and XGBoost pseudo-probability models}
-    \end{center}
-\end{figure}
+In addition, the number of viable output species was reduced from 11255 (the total species present accross all data) down to 1141 (the total species with more than 100 occurrences in the presence/absence data).
+This reduction served to remove several edge cases and focus the models more on likelier species.
 

### 2024-06-23 dataformm.py

@@ -16,12 +16,16 @@ class TestDataset(Dataset):
         self.transform = transform
 
     def __len__(self):
-        return self.pca.shape[0]
+        return self.pca_data.shape[0]
 
     def __getitem__(self, idx):
         survey_id = self.pca_data.index[idx]
-        image_data = self.process_patch(survey_id)
-        sample = [self.pca_data.iloc[idx].tolist(), image_data, self.landsat_data[idx]]
+        # image_data = self.process_patch(survey_id)
+        sample = [
+            self.pca_data.iloc[idx].tolist(), 
+            # image_data, 
+            self.landsat_data[idx]
+        ]
         return sample, survey_id
 
     def process_patch(self, survey_id):
@@ -58,11 +62,11 @@ class TrainDataset(TestDataset):
 
     def __getitem__(self, idx):
         survey_id = self.pca_data.index[idx]
-        image_data = self.process_patch(survey_id)
+        # image_data = self.process_patch(survey_id)
 
         sample = [
             torch.tensor(self.pca_data.iloc[idx].tolist(), dtype=torch.float32),
-            torch.tensor(image_data, dtype=torch.float32),
+            # torch.tensor(image_data, dtype=torch.float32),
             torch.tensor(self.landsat_data[idx], dtype=torch.float32),
         ]
         return (

### 2024-06-23 pipeline/model_management/xgb_manager.py

@@ -16,9 +16,9 @@ class GeoLifeXGB():
 
     __count_params = {
         'device'        : 'cuda',
+        'reg_lambda'    : 10,
         'learning_rate' : 0.1,
         'min_split_loss': 0,
-        'reg_lambda'    : 10,
         'reg_alpha'     : 0, 
     }
 
@@ -46,13 +46,22 @@ class GeoLifeXGB():
         counts = self.count_model.predict(self.test) + addition 
         return counts.round().astype(int)
 
-    def generate_main_model(self):
+    def generate_main_model(self, min_species_occurrence:int = 100):
+        new_yy_cols = None
+        if min_species_occurrence > 100:
+            yy_counts = self.train[self.yy_cols]                               \
+                                            .sum().sort_values(ascending=False)
+            new_yy_cols = yy_counts[:(yy_counts > min_species_occurrence)      \
+                                            .sum()].index
+
         self.main_model = XGBRegressor(**self.__main_params, seed=self.seed)
         combined_train = self.train.drop(columns=self.yy_cols)
         combined_train[
             ["landsat" + str(idx) for idx in range(360)]
         ] = self.landsat
-        self.main_model.fit(combined_train, self.train[self.yy_cols])
+        self.main_model.fit(combined_train, self.train[
+            self.yy_cols if new_yy_cols is None else new_yy_cols
+        ])
         return self.main_model 
     
     def predict_test_scores(self):
@@ -64,13 +73,19 @@ class GeoLifeXGB():
         ] = self.landsat_test 
         return self.main_model.predict(combined_test)
  
-    def __load_data(self):
-        self.train = pk.load(open("data/processed/train.pkl", "rb"))
-        self.test = pk.load(open("data/processed/test.pkl", "rb"))
+    def __load_data(
+            self, 
+            train_path="data/processed/train.pkl", 
+            test_path="data/processed/test.pkl", 
+            train_ls_path="data/processed/train_ls.pkl", 
+            test_ls_path="data/processed/test_ls.pkl"
+        ):
+        self.train = pk.load(open(train_path, "rb"))
+        self.test = pk.load(open(test_path, "rb"))
         self.yy_cols = [col for col in self.train if "speciesId_" in col]
-        self.landsat = pk.load(open("data/processed/train_ls.pkl", "rb"))
+        self.landsat = pk.load(open(train_ls_path, "rb"))
[…]

### 2024-06-23 pipeline/preprocessing/pcamanager.py

@@ -2,6 +2,7 @@ import pandas as pd
 from sklearn.decomposition import PCA
 from numpy import inf, nan, float32, float64
 from warnings import simplefilter
+
 simplefilter(action="ignore", category=pd.errors.PerformanceWarning)
 
 class GeoLifePCA():
